---
title: "LookingatGMSEUS"
author: "Mia Dagati"
date: "2025-09-18"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(sf)
library(tigris)
library(leaflet)
library(plotly)
library(terra)
library(viridis) 
library(patchwork)
library(cowplot)
library(ggrepel)

options(tigris_use_cache = TRUE)

#     Loading data 
array_file <- "C:/Users/miada/Downloads/Energy_Lab/GMSEUS_Arrays_Final_2.1.csv" ## new data set 07/20
panel_file <- "C:/Users/miada/Downloads/Energy_Lab/GMSEUS_Panels_Final_2.0.csv"

array_df <- read_csv(array_file) %>%
  mutate(
    Original_df = "Array",
    capMW       = capMWDC,      # edited due to differing variable names for version 1 and 2 
    capMWest    = capMWDCest    # DC requested for MW cap use
  )

panel_df <- read_csv(panel_file) %>%
  mutate(Original_df = "Panel")
```


## 1. Merge Array + Panel Data

```{r merge-array-panel}
# Replace sentinel -9999 values with NA across panel measurement columns
panel_summary <- panel_df %>%
  mutate(across(c(rowWidth, rowLength, rowAzimuth, rowSpace), ~na_if(., -9999))) %>%
  group_by(arrayID) %>%
  # Flag readings >3 SD from each array's own mean as outliers and set to NA
  # This prevents a single bad reading from skewing the per-array average
  mutate(
    outlier_w = abs(rowWidth  - mean(rowWidth,  na.rm = TRUE)) > 3 * sd(rowWidth,  na.rm = TRUE),
    outlier_l = abs(rowLength - mean(rowLength, na.rm = TRUE)) > 3 * sd(rowLength, na.rm = TRUE),
    outlier_s = abs(rowSpace  - mean(rowSpace,  na.rm = TRUE)) > 3 * sd(rowSpace,  na.rm = TRUE),
    rowWidth  = ifelse(outlier_w, NA_real_, rowWidth),
    rowLength = ifelse(outlier_l, NA_real_, rowLength),
    rowSpace  = ifelse(outlier_s, NA_real_, rowSpace)
  ) %>%
  # Collapse panel-level rows to one summary row per array
  summarise(
    avgRowWidth   = mean(rowWidth,   na.rm = TRUE),
    avgRowLength  = mean(rowLength,  na.rm = TRUE),
    avgRowAzimuth = mean(rowAzimuth, na.rm = TRUE),
    avgRowSpace   = mean(rowSpace,   na.rm = TRUE),
    nRows         = n(),
    .groups       = "drop"
  )

# Replace sentinel -9999 values in array-level data, then join panel summary
array_based_df <- array_df %>%
  mutate(across(c(avgWidth, avgLength, avgSpace, GCR1, GCR2, tilt, instYrEst), ~na_if(., -9999))) %>%
  left_join(panel_summary, by = "arrayID")
```


## 2. Continental US Geometries for Visualizations
#### Used Throughout Analysis 

```{r cont-us-geometry}
# Download 2024 US state boundaries (cartographic, cb = TRUE = simplified coastlines)
#
# Filter out non-contiguous states and territories:
non_cont_US <- c("02","15","72","66","78","69","60")   #places like Alaska, islands etc

cont_US <- states(cb = TRUE, year = 2024) %>%
  st_transform(4326) %>%                  # set to standard lat/lon CRS before filtering
  filter(!STATEFP %in% non_cont_US)       # keep contiguous US only
```

```{r Shared Spatial Setup}
crs_proj <- "EPSG:5070"  # standard CRS value to minimize distortion for contiguous US maps

# Build two versions of the US boundary in the projected CRS:
us_p     <- st_union(st_geometry(cont_US)) |> st_as_sf() |> st_transform(crs_proj)   # single dissolved polygon (used for outer border on maps)
states_p <- st_transform(cont_US, crs_proj)                                          # individual state polygons (used for internal state boundaries)

# Raster and kernel parameters for spatial density estimation
cellsize_km  <- 5    # resolution of the output raster grid (5 km x 5 km cells)
bandwidth_km <- 70   # Gaussian smoothing bandwidth which controls how far influence spreads 

cellsize_m <- cellsize_km  * 1000  # convert to meters to match EPSG:5070 units
bw_m       <- bandwidth_km * 1000

# Create an empty raster template covering the contiguous US extent
r_template <- rast(
  ext        = ext(vect(us_p)),
  resolution = cellsize_m,
  crs        = crs_proj
)

# Build a 2D Gaussian kernel for density smoothing (used later)
rad_cells <- ceiling(bw_m * 3 / cellsize_m)   # defines how many cells out the kernel extends (3 standard deviations)
xy        <- seq(-rad_cells, rad_cells)
g_kernel  <- outer(xy, xy, function(xi, yi)   # g_kernel is normalized so all weights sum to 1 (no artificial inflation of values)
  exp(-(xi^2 + yi^2) * cellsize_m^2 / (2 * bw_m^2)))
g_kernel  <- g_kernel / sum(g_kernel)
```



## 3. Module Type Summary Across US

```{r modtype summary: Goldilocks Explination for CdTe Selection for project}
modType_summary <- array_df %>%
  group_by(modType) %>%
  summarise(
    nArrays    = n(),
    totalCapMW = sum(capMW, na.rm = TRUE),
    .groups    = "drop"
  ) %>%
  mutate(avgCapMW = totalCapMW / nArrays) %>%
  arrange(desc(avgCapMW)) %>%
  mutate(modType = factor(modType, levels = modType))  # preserve sort order

ggplot(modType_summary, aes(x = modType, y = avgCapMW, fill = modType)) +
  geom_col(width = 0.65, show.legend = FALSE) +
  geom_text(aes(label = paste0(round(avgCapMW, 1), " MW\nn = ", nArrays)),
            vjust = -0.5, size = 3.2, color = "grey30") +
  scale_fill_manual(values = c(
    "thin-film" = "navy",
    "c-si"      = "#96CDCD",
    "csp"       = "dodgerblue"
  )) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title    = "Average Capacity per Array by Module Type",
    subtitle = "Sorted by average capacity; labels show avg MW and array count",
    x        = "Module Type",
    y        = "Average Capacity per Array (MW)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title      = element_text(face = "bold", size = 13),
    plot.subtitle   = element_text(size = 10, color = "grey40"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    axis.text.x        = element_text(size = 11)
  )
#table(array_df$modType("a-si", "cdte", "cigs", "thin-film", "thin-film-other"))        # unique values + counts of each
```


## 4. Map — All Arrays (Continental US)

```{r map-all-arrays}
array_based_sf <- array_based_df %>%
  filter(is.finite(longitude), is.finite(latitude)) %>%
  mutate(mod_label = case_when(
    modType == "c-si"      ~ "Crystalline Silicon",
    modType == "cdte" ~ "Thin Film (CdTe)",
    modType == "csp"       ~ "Concentrated Solar Power (CSP)"
  )) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
  st_transform("EPSG:5070")

ggplot() +
  geom_sf(data = states_p, fill = "grey95", color = "black", linewidth = 0.15) +
  geom_sf(data = us_p,     fill = NA,       color = "black", linewidth = 0.40) +
  # Plot non-thin-film arrays first so thin-film renders on top and is not obscured as we want it to be highlighted here
  geom_sf(data = array_based_sf %>% filter(modType != "cdte"),
          aes(color = mod_label), size = 1.5, alpha = 0.6) +
  geom_sf(data = array_based_sf %>% filter(modType == "cdte"),
          aes(color = mod_label), size = 1.5, alpha = 0.8) +
  scale_color_manual(
    values = c("Crystalline Silicon"          = "#96CDCD",
               "Thin Film (CdTe)"                    = "navy",
               "Concentrated Solar Power (CSP)" = "dodgerblue"),
    name = "Module Type"
  ) +
  coord_sf(crs = "EPSG:5070", expand = FALSE) +
  theme_void() +
  theme(
    plot.title      = element_text(face = "bold", size = 13),
    plot.subtitle   = element_text(size = 10, color = "grey40"),
    legend.position = "bottom"
  ) +
  labs(
    title    = "Solar Arrays by Module Type",
    subtitle = "Crystalline silicon, thin-film (CdTe), and CSP installations across the contiguous US"
  )
```


## 5. Thin-Film Module Type Selection and Series Prediction
```{r series-year-identification-and-boundary-highlight-cases}
# Assigns each installation year to its corresponding First Solar module series
# Based on First Solar's known production timeline for Series 2 through Series 7
assign_series <- function(y) {
  case_when(
    y >= 2006 & y <= 2013 ~ "Series 2",
    y >= 2014 & y <= 2016 ~ "Series 3",
    y >= 2017 & y <= 2019 ~ "Series 4",
    y == 2020             ~ "Series 6",
    y >= 2021 & y <= 2022 ~ "Series 6",
    y >= 2023             ~ "Series 7",
    TRUE                  ~ "Unknown"
  )
}

# Flags years that fall near a series transition boundary
# These arrays may contain a mix of two series and are flagged for sensitivity analysis
assign_boundary <- function(y) {
  case_when(
    y %in% c(2012, 2013)       ~ "S2-S3",
    y %in% c(2015, 2016)       ~ "S3-S4",
    y %in% c(2018, 2019, 2020) ~ "S4-S6",
    y == 2022                  ~ "S6-S7",
    TRUE                       ~ NA_character_
  )
}
```


```{r thin-film-series-assignment}
# Filter to thin-film arrays only and append with derived columns
 
thin_film_df <- array_based_df %>%
  filter(modType == "cdte") %>%
  mutate(
    capMW_total      = coalesce(capMW, capMWest),  # uses reported capacity, falling back to estimated if missing
    STATEFP          = str_pad(as.character(STATEFP), 2, pad = "0"), # zero-padded to 2 digits for consistent joining with Census geometries- based on recommendation 
    
    predicted_series = assign_series(instYr),    # First Solar series assigned from installation year
    boundary_case    = assign_boundary(instYr),  # flags arrays installed near a series transition year
    yr_confidence    = case_when(
      abs(instYr - instYrEst) <= 1 ~ "high",     # confidence in instYr based on agreement with instYrEst - need to go back and edit based on new understanding of instYr variable
      abs(instYr - instYrEst) <= 3 ~ "medium",
      TRUE                         ~ "low"
    )
  ) %>%
  # Join state names from Census geometry using zero-padded STATEFP
  left_join(
    cont_US %>% st_set_geometry(NULL) %>%
      transmute(STATEFP = str_pad(as.character(STATEFP), 2, pad = "0"), state_name = NAME),
    by = "STATEFP"
  )

# Quick check on how many arrays fall into each confidence tier
table(thin_film_df$yr_confidence)

# Set series as an ordered factor so plots and tables display in chronological order
series_levels <- c("Series 2", "Series 3", "Series 4", "Series 6", "Series 7")
thin_film_df <- thin_film_df %>%
  mutate(predicted_series = factor(predicted_series, levels = series_levels))
```

## 6. Map — Thin-Film Series (Continental US)

```{r Leaflet: map thin-film series}
# Color palette mapping each First Solar series to a consistent color
# Used across all thin-film maps for visual continuity
series_colors <- c(
  "Series 2" = "lightgreen",
  "Series 3" = "#5AAE61",
  "Series 4" = "dodgerblue",
  "Series 6" = "orchid",
  "Series 7" = "#4B0082"
)
# leaflet-compatible factor palette; brown flags any arrays with NA series (Unknown)
series_palette <- colorFactor(
  palette  = unname(series_colors),
  levels   = names(series_colors),
  na.color = "brown"
)

# Interactive map of thin-film arrays colored by predicted First Solar series
# Popups show full array metadata including year confidence and boundary case status
leaflet(thin_film_df,
        options = leafletOptions(zoomSnap = 0.15, zoomDelta = 0.25, wheelPxPerZoomLevel = 120)) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = cont_US, color = "black", weight = 1, fill = FALSE) %>%
  addCircleMarkers(
    lng   = ~longitude,
    lat   = ~latitude,
    color = ~series_palette(predicted_series),
    popup = ~paste0(
      "<b>Array ID:</b> ",          arrayID,          "<br>",
      "<b>Installation Year:</b> ", instYr,           "<br>",
      "<b>Est. Year (LT):</b> ",    instYrEst,        "<br>",
      "<b>Year Confidence:</b> ",   yr_confidence,    "<br>",
      "<b>Predicted Series:</b> ",  predicted_series, "<br>",
      "<b>Boundary Case:</b> ",     ifelse(is.na(boundary_case), "No", boundary_case), "<br>",
      "<b>Capacity (MW):</b> ",     capMW_total,      "<br>",
      "<b>Coordinates:</b> ",       latitude, ", ", longitude
    ),
    radius = 4, stroke = FALSE, fillOpacity = 1
  ) %>%
  addLegend("bottomright", pal = series_palette, values = ~predicted_series,
            title = "Thin-Film Series") %>%
  # Nudges legend away from the map edge for cleaner rendering in HTML output
  htmlwidgets::onRender("
    function(el, x) {
      $('.leaflet-bottom.leaflet-right').css('bottom','10px').css('right','45px');
    }
  ")
```

```{r Ggplot-static: map thin-film series}
# Convert thin-film arrays to sf and reproject to Albers Equal Area (EPSG:5070)
thin_film_sf <- thin_film_df %>%
  filter(is.finite(longitude), is.finite(latitude)) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
  st_transform("EPSG:5070")

# Static map of thin-film arrays colored by predicted First Solar series
p_series_map <- ggplot() +
  geom_sf(data = states_p, fill = "grey95", color = "black", linewidth = 0.15) +
  geom_sf(data = us_p,     fill = NA,       color = "black", linewidth = 0.40) +
  geom_sf(data = thin_film_sf,
          aes(color = predicted_series), size = 2.5, alpha = 0.9) +
  scale_color_manual(
    values = series_colors,
    name   = "Series\nNumber",
    drop   = FALSE
  ) +
  coord_sf(crs = "EPSG:5070", expand = FALSE) +
  guides(
    color = guide_legend(override.aes = list(size = 3))
  ) +
  theme_void() +
  theme(
    plot.title      = element_text(face = "bold", size = 13),
    plot.subtitle    = element_text(size = 10, color = "grey40"),
    legend.position  = c(0.95, 0.60),   # right side, vertically around Maine
    legend.justification = c(0.5, 1),
    legend.background = element_rect(fill = "white", color = NA),
    legend.key.size  = unit(0.4, "cm"),
    legend.text      = element_text(size = 8),
    legend.title     = element_text(size = 9, face = "bold", hjust = 0.5),
    plot.background   = element_rect(fill = "white", color = NA),
    ) 

p_series_map

ggsave(
  filename = "C:/Users/miada/Downloads/Energy_Lab/Thin-FilmSolarArraysbyPredictedFirstSolarSeries.png",
  plot     = p_series_map,
  width    = 8,
  height   = 5,
  dpi      = 300
)
```
```{r Timeline of all First Solar series installations}
#this will go on top of the figure above, we'll just overlay it. 

all_series_years <- tibble(
  predicted_series = c("Series 2", "Series 3", "Series 4", "Series 6", "Series 7"),
  start_year       = c(2006, 2013, 2016, 2018, 2023),
  end_year         = c(2012, 2015, 2019, 2022, 2025)
)

overlap_window <- tibble(start = 2018, end = 2019)  # overlap of Series 4 (ends 2019) and Series 6 (starts 2018)

# Matches existing series_colors palette (used in the Leaflet map, lines 268-274)
timeline_colors <- c(
  "Series 2" = "lightgreen",
  "Series 3" = "#5AAE61",
  "Series 4" = "dodgerblue",
  "Series 6" = "orchid",
  "Series 7" = "#4B0082"
)

p_timeline_all <- ggplot(all_series_years) +
  geom_rect(data = overlap_window,
            aes(xmin = start, xmax = end, ymin = 0.4, ymax = 5.6),
            fill = "grey80", color = "black", linewidth = 0.8,
            inherit.aes = FALSE) +
  geom_segment(aes(x = start_year, xend = end_year,
                    y = predicted_series, yend = predicted_series,
                    color = predicted_series),
               linewidth = 14, lineend = "round") +
  geom_text(aes(x = (start_year + end_year) / 2, y = predicted_series,
                label = paste0(start_year, "\u2013", end_year)),
            color = "white", fontface = "bold", size = 5) +
  annotate("text", x = 2018.5, y = 5.6,
           label = "S4/S6 overlap\n(2018\u20132019)",
           vjust = -0.3, size = 4.5, fontface = "italic", color = "grey20") +
  scale_color_manual(values = timeline_colors, guide = "none") +
  scale_x_continuous(breaks = seq(2006, 2025, by = 2)) +
  scale_y_discrete(limits = c("Series 2", "Series 3", "Series 4", "Series 6", "Series 7")) +
  labs(x = "Installation Year", y = NULL) +      # title = "Timeline of First Solar Series Installations"
  theme_minimal(base_size = 18) +
  theme(panel.grid.minor  = element_blank(),
        axis.text.y       = element_text(face = "bold", size = 16),
        axis.text.x       = element_text(size = 13),
        axis.title.x      = element_text(size = 15),
        plot.title        = element_text(size = 20, face = "bold"))

p_timeline_all

ggsave(
  filename = "C:/Users/miada/Downloads/Energy_Lab/series_installation_timeline.png",
  plot     = p_timeline_all,
  width    = 14,
  height   = 4,
  dpi      = 300,
  bg       = "white"
)
```

```{r}
kmeans_features_panel <- c("instYr", "latitude", "longitude", "avgWidth", "avgSpace", "GCR1")

# Training set: confident Series 4 (2017) and Series 6 (2021),
# restricted to arrays that actually have avgWidth/avgSpace data
train_df_panel <- thin_film_df %>%
  filter(
    (instYr == 2017, instYr == 2016 & predicted_series == "Series 4") |
    (instYr == 2021, instYr == 2022 & predicted_series == "Series 6")
  ) %>%
  select(arrayID, predicted_series, all_of(kmeans_features_panel)) %>%
  drop_na(all_of(kmeans_features_panel))

table(train_df_panel$predicted_series)

# Scale and fit
train_scaled_panel <- train_df_panel %>%
  select(all_of(kmeans_features_panel)) %>%
  scale()

set.seed(42)
km_fit_panel <- kmeans(train_scaled_panel, centers = 2, nstart = 25)

train_df_panel$cluster <- factor(km_fit_panel$cluster)

table(Actual = train_df_panel$predicted_series, Cluster = train_df_panel$cluster)
```
```{r}
kmeans_features_panel <- c( "latitude", "longitude", "avgWidth", "avgSpace", "GCR1")

# Training set: confident Series 4 (2017) and Series 6 (2021), full sample
train_df_imputed <- thin_film_df %>%
  filter(
    (instYr == 2017 & predicted_series == "Series 4") |
    (instYr == 2021 & predicted_series == "Series 6")
  ) %>%
  select(arrayID, predicted_series, all_of(kmeans_features_panel))

# Check how many are missing before imputing, for transparency
train_df_imputed %>%
  group_by(predicted_series) %>%
  summarise(
    missing_avgWidth = sum(is.na(avgWidth)),
    missing_avgSpace = sum(is.na(avgSpace)),
    n = n()
  )

# Impute avgWidth/avgSpace using each series' own median
# (keeps Series 4 and Series 6 distributions distinct rather than blending them)
train_df_imputed <- train_df_imputed %>%
  group_by(predicted_series) %>%
  mutate(
    avgWidth = ifelse(is.na(avgWidth), median(avgWidth, na.rm = TRUE), avgWidth),
    avgSpace = ifelse(is.na(avgSpace), median(avgSpace, na.rm = TRUE), avgSpace)
  ) %>%
  ungroup() %>%
  drop_na(all_of(kmeans_features_panel))  # safety net in case a whole series is missing

table(train_df_imputed$predicted_series)

# Scale and fit
train_scaled_imputed <- train_df_imputed %>%
  select(all_of(kmeans_features_panel)) %>%
  scale()

set.seed(42)
km_fit_imputed <- kmeans(train_scaled_imputed, centers = 2, nstart = 25)

train_df_imputed$cluster <- factor(km_fit_imputed$cluster)

table(Actual = train_df_imputed$predicted_series, Cluster = train_df_imputed$cluster)
```

```{r}
# Does effInit vary within a single predicted_series across different years?
# If it's a real per-array measurement, we'd expect some spread.
# If it's a lookup constant by series, every year within that series will be identical.

thin_film_df %>%
  filter(modType == "cdte", instYr >= 2016, instYr <= 2021) %>%
  group_by(instYr) %>%
  summarise(
    mean_eff = mean(effInit, na.rm = TRUE),
    sd_eff   = sd(effInit, na.rm = TRUE),
    n        = n(),
    .groups  = "drop"
  ) %>%
  arrange(instYr)
thin_film_df %>%
  filter(instYr >= 2016, instYr <= 2022) %>%
  group_by(instYr) %>%
  summarise(
    total_arrays         = n(),
    missing_avgWidth     = sum(is.na(avgWidth)),
    missing_avgSpace     = sum(is.na(avgSpace)),
    missing_either       = sum(is.na(avgWidth) | is.na(avgSpace)),
    pct_missing_either   = round(100 * missing_either / total_arrays, 1),
    .groups = "drop"
  ) %>%
  arrange(instYr)
```
```{r}
thin_film_df %>%
  filter(instYr >= 2016, instYr <= 2022) %>%
  summarise(
    total_arrays       = n(),
    missing_avgWidth   = sum(is.na(avgWidth)),
    missing_avgSpace   = sum(is.na(avgSpace)),
    missing_either     = sum(is.na(avgWidth) | is.na(avgSpace)),
    pct_missing_either = round(100 * missing_either / total_arrays, 1)
  )
```


```{r}
thin_film_df %>%
  filter(
    (instYr == 2017 & predicted_series == "Series 4") |
    (instYr == 2021 & predicted_series == "Series 6")
  ) %>%
  summarise(
    missing_avgWidth    = sum(is.na(avgWidth)),
    missing_avgRowWidth = sum(is.na(avgRowWidth)),
    missing_avgSpace    = sum(is.na(avgSpace)),
    missing_avgRowSpace = sum(is.na(avgRowSpace))
  )
```


```{r}
# --- Deploy trained clusters onto the boundary years (2018, 2019, 2020) ---
boundary_df <- thin_film_df %>%
  filter(boundary_case == "S4-S6") %>%
  select(arrayID, instYr, predicted_series, boundary_case, all_of(kmeans_features)) %>%
  drop_na(all_of(kmeans_features))

# Apply the SAME scaling (mean/sd from training data, not refit on boundary data,
# so boundary points land in the same coordinate space as the trained centers)
center_means <- attr(scale(train_df %>% select(all_of(kmeans_features))), "scaled:center")
center_sds   <- attr(scale(train_df %>% select(all_of(kmeans_features))), "scaled:scale")

boundary_scaled <- boundary_df %>%
  mutate(across(all_of(kmeans_features),
                ~(. - center_means[cur_column()]) / center_sds[cur_column()]))

# kmeans() has no native predict() method, so assign each boundary point
# to whichever of the 2 trained centers it's closest to (Euclidean distance)
assign_cluster <- function(x, centers) {
  d <- sapply(1:nrow(centers), function(i) sum((x - centers[i, ])^2))
  which.min(d)
}

boundary_scaled$cluster <- apply(
  boundary_scaled %>% select(all_of(kmeans_features)), 1,
  assign_cluster, centers = km_fit$centers
)
boundary_scaled$cluster <- factor(boundary_scaled$cluster)

boundary_df <- boundary_df %>%
  bind_cols(cluster = boundary_scaled$cluster) %>%
  left_join(cluster_to_series, by = "cluster") %>%
  rename(kmeans_predicted_series = cluster_series)

# Compare original year-based guess vs. k-means-refined guess for boundary arrays
boundary_df %>%
  count(instYr, predicted_series, kmeans_predicted_series) %>%
  arrange(instYr)
```
```{r}
# Join state_name back onto train_df (it was dropped in the select() step)
train_df_states <- train_df %>%
  left_join(
    thin_film_df %>% select(arrayID, state_name),
    by = "arrayID"
  )

# Focus on Cluster 2, where Series 4 and Series 6 overlap
cluster2_breakdown <- train_df_states %>%
  filter(cluster == 2) %>%
  count(predicted_series, state_name) %>%
  arrange(desc(n))

print(cluster2_breakdown, n = 50)

# Side-by-side: which states show up for S4-in-cluster2 vs S6-in-cluster2
cat("\n=== States for Series 4 arrays landing in Cluster 2 ===\n")
train_df_states %>%
  filter(cluster == 2, predicted_series == "Series 4") %>%
  count(state_name, sort = TRUE) %>%
  print(n = 20)

cat("\n=== States for Series 6 arrays landing in Cluster 2 ===\n")
train_df_states %>%
  filter(cluster == 2, predicted_series == "Series 6") %>%
  count(state_name, sort = TRUE) %>%
  print(n = 20)

# Overlap: states where BOTH series show up in cluster 2 (the truly ambiguous ones)
overlap_states <- intersect(
  train_df_states %>% filter(cluster == 2, predicted_series == "Series 4") %>% pull(state_name),
  train_df_states %>% filter(cluster == 2, predicted_series == "Series 6") %>% pull(state_name)
)
cat("\n=== States where Series 4 AND Series 6 both land in Cluster 2 ===\n")
print(overlap_states)
```





## 7. Module Count Estimation

```{r thin-film module count estimation}
# Series-level fallback module areas (m²) from First Solar spec sheets
# W = short edge, L = long edge per GM-SEUS convention
# Module area = W * L
module_sizes <- c(
  "Series 2" = 0.72,  # 1.200m x 0.600m
  "Series 3" = 0.72,  # 1.200m x 0.600m
  "Series 4" = 0.72,  # 1.200m x 0.600m
  "Series 6" = 2.47,  # 1.232m x 2.009m
  "Series 7" = 2.80   # 1.216m x 2.300m
)

# Reshape to a dataframe for joining onto thin_film_df by predicted_series
module_sizes_df <- tibble(
  predicted_series = names(module_sizes),
  module_size_m2   = module_sizes
)

thin_film_df <- thin_film_df %>%
  mutate(predicted_series = as.character(predicted_series)) %>%  # drop factor before join
  left_join(module_sizes_df, by = "predicted_series") %>%
  mutate(
    # Three-tier module area hierarchy:
    # 1) Panel geometry  — outlier-cleaned avgRowWidth/Length from panel join
    # 2) Array geometry  — avgWidth/Length from array file
    # 3) Series lookup   — fixed size from spec sheet (last resort)
    effective_module_w    = coalesce(avgRowWidth,  avgWidth),
    effective_module_l    = coalesce(avgRowLength, avgLength),
    effective_module_area = effective_module_w * effective_module_l,
    module_area_used = coalesce(
      ifelse(is.finite(effective_module_area) & effective_module_area > 0,
             effective_module_area, NA_real_),
      module_size_m2
    ),
    # Divide total row area by per-module area to estimate module count per array
    num_modules = round(totRowArea / module_area_used),
    # Track which tier was used for each array
    modules_source = case_when(
      !is.na(avgRowWidth)             ~ "panel_geometry",
      !is.na(avgWidth) & avgWidth > 0 ~ "array_geometry",
      TRUE                            ~ "series_lookup"
    )
  )

# Breakdown of how many arrays used each method
table(thin_film_df$modules_source)
head(thin_film_df$num_modules)
nrow(thin_film_df)
```
## 8. Material Estimation per Array

```{r material-estimation}
thin_film_df <- thin_film_df %>%
  select(-any_of(c("CdTe_g","Aluminum_kg","Steel_kg","EVA_kg","Copper_g",
                    "GlassMass_kg","Other_kg","TotalWeight_kg","CalcTotalMass_kg")))
# Material composition per module by series
# Source: First Solar product datasheets
# CdTe and Copper in grams/module; all others in kg/module
material_per_module <- tibble(
  predicted_series = c("Series 2", "Series 3", "Series 4", "Series 6", "Series 7"),
  CdTe_g           = c(13,    13,    11,    36,    41   ),
  Aluminum_kg      = c(0,     0,     0,     3.93,  0    ),  # was 4.2
  Steel_kg         = c(0,     0,     0,     0,     6.5  ),  # was 6.97
  EVA_kg           = c(0.28,  0.28,  0.28,  0.97,  1.11 ),
  Copper_g         = c(43.7,  27.3,  27.3,  16.1,  14.6 ),
  GlassMass_kg      = c(11.52, 11.52, 11.52, 30.875, 35.00),
  Other_kg         = c(0.3,   0.3,   0.3,   0.3,   0.3  ),
  TotalWeight_kg   = c(12,    12,    12,    34.5,  39.7 ),
  CalcTotalMass_kg  = c(12.1567, 12.1403, 12.1383, 36.1271, 42.9656)  # was 36.3971, 43.4356
)

thin_film_df <- thin_film_df %>%
  left_join(material_per_module, by = "predicted_series") %>%
  mutate(
    total_CdTe_g          = CdTe_g         * num_modules,
    total_Aluminum_kg     = Aluminum_kg    * num_modules,
    total_Steel_kg        = Steel_kg       * num_modules,
    total_EVA_kg          = EVA_kg         * num_modules,
    total_Copper_g        = Copper_g       * num_modules,
    total_Other_kg        = Other_kg       * num_modules,
    total_Array_Weight_kg = TotalWeight_kg * num_modules,
    total_GlassMass_kg    = GlassMass_kg     * num_modules,
    total_CalcMass_kg     = CalcTotalMass_kg * num_modules,
    total_CdTe_metton     = total_CdTe_g   / 1e6,  # g → metric tons
    total_Copper_metton   = total_Copper_g / 1e6   # g → metric tons
  )
```

## 9. Series Distribution Summary

```{r series-distribution overview}
# NOTE: material columns (total_GlassMass_kg, total_CalcMass_kg) are joined in
# Section 9 (material-estimation). Run that chunk first if executing out of order.
series_distribution <- thin_film_df %>%
  group_by(predicted_series) %>%
  summarise(
    num_arrays        = n(),
    total_modules     = sum(num_modules,         na.rm = TRUE),
    total_Glass_t     = sum(total_GlassMass_kg,  na.rm = TRUE) / 1e3,
    total_CalcMass_t  = sum(total_CalcMass_kg,   na.rm = TRUE) / 1e3,
    .groups = "drop"
  ) %>%
  mutate(
    percent_arrays    = round(100 * num_arrays       / sum(num_arrays),       1),
    percent_modules   = round(100 * total_modules    / sum(total_modules),    1),
    pct_Glass         = round(100 * total_Glass_t    / sum(total_Glass_t),    1),
    pct_CalcMass      = round(100 * total_CalcMass_t / sum(total_CalcMass_t), 1)
  ) %>%
  arrange(desc(total_modules))

series_distribution
```


## 10. State-Level Material Aggregation

```{r state-materials}
state_materials <- thin_film_df %>%
  mutate(STATEFP = str_pad(as.character(STATEFP), 2, pad = "0")) %>%
  filter(!STATEFP %in% non_cont_US) %>%
  group_by(STATEFP) %>%
  summarise(
    CdTe_t     = sum(total_CdTe_g,       na.rm = TRUE) / 1e6,  #  g -> t
    Aluminum_t = sum(total_Aluminum_kg,  na.rm = TRUE) / 1e3,  # kg -> t
    Steel_t    = sum(total_Steel_kg,     na.rm = TRUE) / 1e3,  # kg  
    EVA_t      = sum(total_EVA_kg,       na.rm = TRUE) / 1e3,  # kg 
    Copper_t   = sum(total_Copper_g,     na.rm = TRUE) / 1e6,  #  g 
    Glass_t    = sum(total_GlassMass_kg, na.rm = TRUE) / 1e3,  # kg
    Other_t    = sum(total_Other_kg,     na.rm = TRUE) / 1e3,  # kg 
    CalcMass_t = sum(total_CalcMass_kg,  na.rm = TRUE) / 1e3,  # kg
    .groups = "drop"
  )

us_materials <- cont_US %>%
  left_join(state_materials, by = "STATEFP")
```


## 11. Choropleth Maps by Material

```{r choropleth-maps}
# Reusable function to generate a state-level choropleth for a given material column
# Expects us_materials to have one row per state with pre-aggregated tonnage columns
create_material_map <- function(material_col, material_name) {
  pal <- colorNumeric(palette = "plasma", domain = us_materials[[material_col]], na.color = "grey90")

  leaflet(us_materials) %>%
    addProviderTiles("CartoDB.Positron") %>%
    addPolygons(
      fillColor = ~pal(get(material_col)),
      color = "black", weight = 1, fillOpacity = 0.8,
      label = ~paste0(NAME, ": ", round(get(material_col), 1), " t"),
      highlightOptions = highlightOptions(weight = 2, color = "red", bringToFront = TRUE)
    ) %>%
    addLegend("bottomright", pal = pal, values = ~get(material_col),
              title = paste("Total", material_name, "(t)"), opacity = 0.7)
}

create_material_map("CdTe_t",     "Cadmium Telluride")
create_material_map("Aluminum_t", "Aluminum")
create_material_map("Steel_t",    "Steel")
create_material_map("EVA_t",      "EVA Encapsulant")
create_material_map("Copper_t",   "Copper")
create_material_map("Other_t",    "Other Materials")
create_material_map("Glass_t",    "Glass Mass")
create_material_map("CalcMass_t", "Total Calculated Mass")
```


## 12. Material Density Maps

```{r density-map-function}
# Reusable weighted KDE density map function for any numeric column in thin_film_df
# When material_col is NULL, produces a simple point density map (all weights = 1)
# When material_col is provided, each point is weighted by that column's value (e.g. tonnes of CdTe)
create_material_density_map <- function(
    material_col       = NULL,         # column to weight by; NULL = unweighted point density
    material_name      = material_col, # legend title; defaults to column name if not specified
    df                 = thin_film_df,
    lon_col            = "longitude",
    lat_col            = "latitude",
    cellsize_km        = 10,           # raster grid resolution in km
    bandwidth_km       = 90,           # Gaussian smoothing bandwidth in km
    palette_option     = "magma",      # any viridis palette option (magma, plasma, viridis etc.)
    raster_interpolate = FALSE) {      # bilinear interpolation on raster; FALSE = sharp grid cells

  # Select only the columns needed and drop rows with non-finite coordinates or weights
  needed <- c(lon_col, lat_col)
  if (!is.null(material_col)) needed <- c(needed, material_col)
  d <- df[, needed, drop = FALSE]
  d <- d[is.finite(d[[lon_col]]) & is.finite(d[[lat_col]]), , drop = FALSE]
  if (!is.null(material_col)) d <- d[is.finite(d[[material_col]]), , drop = FALSE]

  # Set weights: uniform (1) for point density, or the material column values for weighted KDE
  w <- if (is.null(material_col)) rep(1, nrow(d)) else d[[material_col]]

  # Project points to Albers Equal Area (EPSG:5070) to match raster CRS
  pts <- st_as_sf(d, coords = c(lon_col, lat_col), crs = 4326, remove = FALSE) |>
    st_transform(crs_proj)

  # Build empty raster grid over the contiguous US extent at the specified resolution
  cellsize_m <- cellsize_km * 1000
  r          <- rast(ext = ext(vect(us_p)), resolution = cellsize_m, crs = crs_proj)

  # Rasterize points by summing weights into each grid cell
  pts_spat   <- vect(pts)
  pts_spat$w <- w
  r_sum      <- rasterize(pts_spat, r, field = "w", fun = "sum", background = 0)

  # Build and apply a 2D Gaussian kernel for spatial smoothing
  # Kernel extends 3 standard deviations out from each cell (captures ~99.7% of influence)
  # Normalized so weights sum to 1 — preserves total mass across the raster
  bw_m      <- bandwidth_km * 1000
  rad_cells <- ceiling(bw_m * 3 / cellsize_m)
  xy        <- seq(-rad_cells, rad_cells)
  g         <- outer(xy, xy, function(xi, yi) exp(-(xi^2 + yi^2) * cellsize_m^2 / (2 * bw_m^2)))
  g         <- g / sum(g)
  r_smooth  <- focal(r_sum, w = g, fun = sum, na.policy = "omit", fillvalue = 0)

  # Clip smoothed raster to US boundary and convert to dataframe for ggplot
  r_smooth    <- mask(r_smooth, vect(us_p))
  df_r        <- as.data.frame(r_smooth, xy = TRUE, na.rm = TRUE)
  names(df_r) <- c("x", "y", "z")

  ggplot() +
    geom_raster(data = df_r, aes(x = x, y = y, fill = z), interpolate = raster_interpolate) +
    geom_sf(data = states_p, fill = NA, color = "black", linewidth = 0.15) +
    geom_sf(data = us_p,     fill = NA, color = "black", linewidth = 0.40) +
    coord_sf(crs = crs_proj, expand = FALSE) +
    scale_fill_viridis(option = palette_option, direction = -1,
                       name = ifelse(is.null(material_name), "Density", material_name)) +
    theme_void() +
    theme(legend.position = "right") +
    labs(
      title    = "Density Map — Continental United States",
      subtitle = if (!is.null(material_col)) paste("Weighted by", material_col) else "Point density"
    )
}
```

```{r run-density-maps}
create_material_density_map(
  material_col  = "total_CdTe_g",
  material_name = "CdTe (g)",
  cellsize_km   = 5,
  bandwidth_km  = 70
)

create_material_density_map(
  material_col  = "total_Steel_kg",
  material_name = "Steel (kg)",
  cellsize_km   = 5,
  bandwidth_km  = 70
)

create_material_density_map(
  material_col  = "total_GlassMass_kg",
  material_name = "Glass Mass (kg)",
  cellsize_km   = 5,
  bandwidth_km  = 70
)

create_material_density_map(
  material_col  = "total_CalcMass_kg",
  material_name = "Total Calculated Mass (kg)",
  cellsize_km   = 5,
  bandwidth_km  = 70
)

```

```{r Mapping Materials Normalizing for porportional Density masking to US grid}
# Materials to map
# Each entry defines the source column, display name, and unit scale factor
# scale = 1/1000 converts grams to kg for CdTe and Copper so all materials are comparable
materials <- list(
  list(col = "total_CdTe_g",      name = "Cadmium Telluride", scale = 1/1000), # g -> kg
  list(col = "total_Aluminum_kg", name = "Aluminum",          scale = 1),
  list(col = "total_Steel_kg",    name = "Steel",             scale = 1),
  list(col = "total_EVA_kg",      name = "EVA Encapsulant",   scale = 1),
  list(col = "total_Copper_g",    name = "Copper",            scale = 1/1000), # g -> kg
  list(col = "total_Other_kg",    name = "Other Materials",   scale = 1),
  list(col = "total_GlassMass_kg", name = "Glass Mass",           scale = 1),
  list(col = "total_CalcMass_kg",  name = "Total Calculated Mass", scale = 1)
)

# Compute proportional density raster for each material 
# Normalizing by US total converts raw material mass to proportion of national total
# This makes materials with very different units/magnitudes visually comparable across facets
material_density_df <- purrr::map_dfr(materials, function(mat) {

  # Subset to needed columns and drop any rows with non-finite values
  d <- thin_film_df[, c("longitude", "latitude", mat$col), drop = FALSE]
  d <- d[is.finite(d$longitude) & is.finite(d$latitude) & is.finite(d[[mat$col]]), ]

  # Normalize each array's value to its share of the US total for this material
  us_total     <- sum(d[[mat$col]], na.rm = TRUE)
  d[[mat$col]] <- d[[mat$col]] / us_total

  # Project points and assign normalized weights for rasterization
  pts <- st_as_sf(d, coords = c("longitude", "latitude"), crs = 4326) |>
    st_transform(crs_proj)
  pts_spat   <- vect(pts)
  pts_spat$w <- d[[mat$col]]

  # Rasterize by summing weights per cell, then smooth with the pre-built Gaussian kernel
  # r_template and g_kernel are defined in the Shared Spatial Setup chunk
  r_sum    <- rasterize(pts_spat, r_template, field = "w", fun = "sum", background = 0)
  r_smooth <- focal(r_sum, w = g_kernel, fun = sum, na.policy = "omit", fillvalue = 0)
  r_smooth <- mask(r_smooth, vect(us_p))

  # Convert raster to dataframe and tag with material name for faceting
  df_r        <- as.data.frame(r_smooth, xy = TRUE, na.rm = TRUE)
  names(df_r) <- c("x", "y", "density")
  df_r$material <- mat$name
  df_r
})

# Full grid + mask to US boundary
# Not every material has a density value at every grid cell after masking
# Expanding to a full grid and filling missing cells with a near-zero value (1e-9)
# ensures consistent facet dimensions and prevents blank panels in the plot
full_grid <- expand.grid(
  x        = unique(material_density_df$x),
  y        = unique(material_density_df$y),
  material = unique(material_density_df$material)
)

material_density_df <- full_grid %>%
  left_join(material_density_df, by = c("x", "y", "material")) %>%
  mutate(density = replace_na(density, 1e-9))  # near-zero rather than NA so color scale renders correctly

# Re-apply US boundary mask — cells outside the contiguous US are set to NA so they render as blank
grid_sf <- st_as_sf(material_density_df, coords = c("x", "y"), crs = crs_proj)
in_us   <- st_intersects(grid_sf, us_p, sparse = FALSE)[, 1]
material_density_df <- material_density_df %>%
  mutate(density = ifelse(in_us, density, NA))
```

```{r Faceted density plot}

# Small offset added to all density values before log transform
# log10(0) is undefined and log10(1e-9) creates extreme outliers in the color scale
# 1e-6 is below any real signal so it acts as a visual floor without distorting the map
material_density_df <- material_density_df %>%
  mutate(density = density + 1e-6)

# Set facet order — CdTe first as the primary material of interest,
# followed by encapsulant and structural materials
material_density_df$material <- factor(material_density_df$material, levels = c(
  "Cadmium Telluride",
  "EVA Encapsulant",
  "Aluminum",
  "Steel",
  "Copper",
  "Glass Mass",
  "Other Materials"
))

ggplot() +
  # us_p drawn first so state lines render on top of the raster
  geom_sf(data = us_p, fill = NA, color = "black", linewidth = 0.25) +
  geom_raster(data = material_density_df,
              aes(x = x, y = y, fill = density)) +
  geom_sf(data = states_p, fill = NA, color = "white", linewidth = 0.05) +  # white lines separate states without overpowering the fill
  coord_sf(crs = crs_proj, expand = FALSE) +
  scale_fill_viridis(
    option    = "magma",
    direction = 1,
    trans     = "log10",        # log scale compresses the wide dynamic range of density values
    breaks    = c(1e-6, 1e-5, 1e-4),
    labels    = c("1e-6", "1e-5", "1e-4"),
    na.value  = "white",        # NA cells (outside US boundary) render as white
    name      = "Proportion\nof U.S. Total"
  ) +
  facet_wrap(~material, ncol = 2) +
  theme_void() +
  theme(
    plot.background   = element_rect(fill = NA, color = NA),
    strip.background  = element_blank(),
    strip.text        = element_text(face = "bold", size = 11),
    panel.spacing     = unit(0.75, "lines"),
    legend.position   = "right",
    legend.title      = element_text(size = 9, face = "bold"),
    legend.key.height = unit(1.2, "cm"),
    plot.title        = element_text(size = 13, face = "bold"),
    plot.subtitle     = element_text(size = 9,  color = "grey40")
  ) +
  labs(
    title    = "Geographic Distribution of Thin-Film Solar Array Materials by Type",
    subtitle = "Material density estimated across U.S. states (normalized to proportion)"
  )
```



```{r Building Series Density df }
# Combined density map showing all series with distinct colors
# Each series plotted as its own layer so colors stay distinct
library(terra)
library(viridis)

# Build a smoothed raster for each series
smooth_series <- function(series_name) {
  pts <- thin_film_df %>%
    filter(predicted_series == series_name,
           is.finite(longitude), is.finite(latitude)) %>%
    st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
    st_transform(crs_proj)

  if (nrow(pts) == 0) return(NULL)

  pts_spat   <- vect(pts)
  pts_spat$w <- 1
  r_sum      <- rasterize(pts_spat, r_template, field = "w", fun = "sum", background = 0)

  r_smooth <- focal(r_sum, w = g_kernel, fun = sum, na.policy = "omit", fillvalue = 0)
  r_smooth <- mask(r_smooth, vect(us_p))

  df <- as.data.frame(r_smooth, xy = TRUE, na.rm = TRUE)
  names(df) <- c("x", "y", "density")
  df$series <- series_name
  df
}

series_density_df <- bind_rows(lapply(
  c("Series 2", "Series 3", "Series 4", "Series 6", "Series 7"),
  smooth_series
))

full_grid <- expand.grid(
  x      = unique(series_density_df$x),
  y      = unique(series_density_df$y),
  series = unique(series_density_df$series)
)

# Left join so missing cells get NA, then fill with near-zero
series_density_df <- full_grid %>%
  left_join(series_density_df, by = c("x", "y", "series")) %>%
  mutate(density = replace_na(density, 1e-6))

grid_sf <- st_as_sf(series_density_df, coords = c("x", "y"), crs = crs_proj)

# Mask to US outline — points outside get NA so when we map it there isnt blank color spaces
in_us <- st_intersects(grid_sf, us_p, sparse = FALSE)[, 1]

series_density_df <- series_density_df %>%
  mutate(density = ifelse(in_us, density, NA))
```


```{r Ggplot of Density maps of Distribution of Array}
series_density_df <- series_density_df %>%
  mutate(density = density + 1e-6)
ggplot() +
  geom_sf(data = us_p,     fill = NA,       color = "black", linewidth = 0.25) +
  geom_raster(data = series_density_df,
              aes(x = x, y = y, fill = density)) +
  geom_sf(data = states_p, fill = NA, color = "white", linewidth = 0.1) +
  coord_sf(crs = crs_proj, expand = FALSE) +
  scale_fill_viridis(
  option    = "magma",
  direction = 1,
  na.value  = "white",
  name      = "Array Density\n (log scale)",
  trans     = "log10",
  breaks    = c(0.0001, 0.001, 0.01),
  labels    = c("0.0001", "0.001", "0.01")
  )+
  facet_wrap(~series, ncol = 2) +
  theme_void() +
  theme(
    
    plot.background  = element_rect(fill = NA, color = NA),
    
    strip.background = element_blank(),
    strip.text       = element_text(face = "bold", size = 11),
    
    panel.spacing    = unit(0.75, "lines"),
    
    legend.position  = "right",
    legend.title     = element_text(size = 9, face = "bold"),
    legend.key.height = unit(1.2, "cm"),
    
    plot.title       = element_text(size = 13, face = "bold"),
    plot.subtitle    = element_text(size = 9,  color = "grey40")
  ) +
  labs(
    title    = "Geographic Distribution of Thin-Film Solar Arrays by Product Series",
    subtitle = "Array density estimated across U.S. states (log scale)"
  )

```

## 5b. Bubble Map — Installation Year and Capacity (Static)
```{r bubble-map-year-capacity CONTINUOUS}
year_counts <- thin_film_df %>%
  filter(is.finite(instYr)) %>%
  group_by(instYr) %>%
  summarise(n = n(), .groups = "drop")

inset_bar <- ggplot(year_counts, aes(x = instYr, y = n, fill = instYr)) +
  geom_col(width = 0.8, color = "black", linewidth = 0.3) +   # black outline on each bar
  scale_fill_viridis_c(option = "magma", guide = "none") +
  scale_x_continuous(breaks = seq(min(year_counts$instYr), max(year_counts$instYr), by = 2)) +
  labs(x = NULL, y = "Installed Arrays") +
  theme_minimal(base_size = 9) +
  theme(
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    plot.background    = element_rect(fill = "white", color = NA),  # border removed (was "grey70")
    axis.text.x        = element_text(angle = 45, hjust = 1, size = 9),
    axis.text.y        = element_text(size = 9)
  )

thin_film_sf <- thin_film_df %>%
  filter(is.finite(longitude), is.finite(latitude),
         is.finite(instYr),    is.finite(capMW_total)) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
  st_transform("EPSG:5070")

us_bbox     <- st_bbox(us_p)
inset_x_min <- us_bbox["xmin"] - (us_bbox["xmax"] - us_bbox["xmin"]) * 0.03  # negative pushes left of map edge
inset_x_max <- us_bbox["xmin"] + (us_bbox["xmax"] - us_bbox["xmin"]) * 0.29  # keep same width
inset_y_min <- us_bbox["ymin"]
inset_y_max <- us_bbox["ymin"] + (us_bbox["ymax"] - us_bbox["ymin"]) * 0.22

ggplot() +
  geom_sf(data = states_p, fill = "grey80", color = "black", linewidth = 0.15) +
  geom_sf(data = us_p,     fill = NA,       color = "black", linewidth = 0.40) +
  geom_sf(data = thin_film_sf,
          aes(size = capMW_total, color = instYr),
          alpha  = 0.5,
        stroke = 0) +
  annotation_custom(
    grob = ggplotGrob(inset_bar),
    xmin = inset_x_min, xmax = inset_x_max,
    ymin = inset_y_min, ymax = inset_y_max
  ) +
  scale_size_continuous(
    range  = c(1, 20),
    name   = "Capacity (MW)",
    breaks = c(1, 10, 50, 100, 400, 600),
    labels = function(x) paste0(round(x, 0))
  ) +
  scale_color_viridis_c(option = "magma", name = "Installation
Year") +
  coord_sf(crs = "EPSG:5070", expand = FALSE) +
  theme_void() +
  theme(
    plot.title      = element_text(face = "bold", size = 13),
    plot.subtitle   = element_text(size = 10, color = "grey40"),
    legend.position = c(0.96, 0.33),  # x=0.92 nudges left from right edge, y=0.5 centers vertically
    legend.text     = element_text(size = 11),
    legend.title    = element_text(size = 12, face = "bold"),
    plot.margin     = margin(t = 5, r = 5, b = 5, l = 40)
  )
  

ggsave(
  filename = "C:/Users/miada/Downloads/Energy_Lab/bubble_map_year_capacity.png",
  width    = 14,
  height   = 8,
  dpi      = 400,
  bg       = "white"
)
```

```{r bubble-map-year-capacity BINNED}

# Summarise annual installation counts for the inset bar chart
year_counts <- thin_film_df %>%
  filter(is.finite(instYr)) %>%
  group_by(instYr) %>%
  summarise(n = n(), .groups = "drop")

# Build inset bar chart — colors match the magma palette used in the bubble map
inset_bar <- ggplot(year_counts, aes(x = instYr, y = n, fill = instYr)) +
  geom_col(width = 0.8) +
  scale_fill_viridis_c(option = "magma", guide = "none") +  # same palette as bubble map, no legend
  scale_x_continuous(breaks = seq(min(year_counts$instYr), max(year_counts$instYr), by = 2)) +
  labs(x = NULL, y = "Arrays") +
  theme_minimal(base_size = 7) +
  theme(
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    plot.background    = element_rect(fill = "white", color = "grey70", linewidth = 0.3),
    axis.text.x        = element_text(angle = 45, hjust = 1, size = 6),
    axis.text.y        = element_text(size = 6)
  )

# Convert thin-film arrays to sf and reproject to Albers Equal Area (EPSG:5070)
thin_film_sf <- thin_film_df %>%
  filter(is.finite(longitude), is.finite(latitude),
         is.finite(instYr),    is.finite(capMW_total)) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
  st_transform("EPSG:5070")

# Get the projected coordinate extent of the US for inset positioning
us_bbox <- st_bbox(us_p)

# Place inset in bottom-left — below California, above the southern border
# Adjust xmin/xmax/ymin/ymax if the inset needs to shift
inset_x_min <- us_bbox["xmin"]
inset_x_max <- us_bbox["xmin"] + (us_bbox["xmax"] - us_bbox["xmin"]) * 0.28
inset_y_min <- us_bbox["ymin"]
inset_y_max <- us_bbox["ymin"] + (us_bbox["ymax"] - us_bbox["ymin"]) * 0.22

ggplot() +
  geom_sf(data = states_p, fill = "grey80", color = "black", linewidth = 0.15) +
  geom_sf(data = us_p,     fill = NA,       color = "black", linewidth = 0.40) +
  # Bubble size = capacity, color = installation year — both encoded for emphasis
  geom_sf(data = thin_film_sf,
          aes(size = capMW_total, color = instYr),
          alpha = 0.5) +
  # Inset bar chart positioned in bottom-left corner of the projected coordinate space
  annotation_custom(
    grob = ggplotGrob(inset_bar),
    xmin = inset_x_min, xmax = inset_x_max,
    ymin = inset_y_min, ymax = inset_y_max
  ) +
  scale_size_binned(
  breaks = c(1, 10, 25, 100, 300),
  range  = c(1, 20),
  name   = "Capacity (MW)",
  labels = c("1-10", "10-25", "25-100", "100-300", "300+"),
  guide  = guide_bins(
    axis       = FALSE,
    show.limits = FALSE,
    keywidth   = unit(0.5, "cm"),
    keyheight  = unit(0.5, "cm")
  )
) +
  scale_color_viridis_c(option = "magma", name = "Installation Year") +
  coord_sf(crs = "EPSG:5070", expand = FALSE) +
  theme_void() +
  theme(
    plot.title      = element_text(face = "bold", size = 13),
    plot.subtitle   = element_text(size = 10, color = "grey40"),
    legend.position = "right"
  ) +
  labs(
    title    = "Thin-Film Solar Arrays by Installation Year and Capacity",
    subtitle = "Bubble size = capacity (MW), color = installation year; inset = annual installation count"
  )
```


```{r line-graph-avg-median-cum}
line_df <- thin_film_df %>%
  filter(is.finite(instYr), is.finite(capMW_total)) %>%
  group_by(instYr) %>%
  summarise(
    mean_cap     = mean(capMW_total,   na.rm = TRUE),
    median_cap   = median(capMW_total, na.rm = TRUE),
    annual_count = n(),
    .groups = "drop"
  ) %>%
  arrange(instYr) %>%
  mutate(cumulative_count = cumsum(annual_count))

scale_factor <- max(c(line_df$mean_cap, line_df$median_cap)) / max(line_df$cumulative_count)

ggplot(line_df, aes(x = instYr)) +
  geom_line(aes(y = mean_cap,   color = "Mean Capacity (MW)"),   linewidth = 1.2) +
  geom_point(aes(y = mean_cap,  color = "Mean Capacity (MW)"),   size = 2.5, shape = 16) +
  geom_line(aes(y = median_cap, color = "Median Capacity (MW)"), linewidth = 1.2, linetype = "dashed") +
  geom_point(aes(y = median_cap,color = "Median Capacity (MW)"), size = 3.5, shape = 18) +
  geom_line(aes(y = cumulative_count * scale_factor,  color = "Cumulative Installations"), linewidth = 1.2) +
  geom_point(aes(y = cumulative_count * scale_factor, color = "Cumulative Installations"), size = 2.5) +
  scale_y_continuous(
    name     = "Capacity per Array (MW)",
    sec.axis = sec_axis(~ . / scale_factor, name = "Cumulative Array Count")
  ) +
  scale_x_continuous(breaks = seq(min(line_df$instYr), max(line_df$instYr), by = 2)) +
  scale_color_manual(
    values = c("Mean Capacity (MW)"       = "green4",
               "Median Capacity (MW)"     = "chartreuse3",
               "Cumulative Installations" = "dodgerblue4"),
    name   = NULL
  ) +
  labs(
    title    = "Mean and Median Array Capacity Against Cumulative Installations",
    subtitle = "Mean (solid) and median (dashed) capacity per array vs cumulative array count by year",
    x        = "Installation Year"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title         = element_text(face = "bold", size = 13),
    plot.subtitle      = element_text(size = 10, color = "grey40"),
    legend.position    = "top",
    panel.grid.minor   = element_blank(),
    axis.title.y.right = element_text(color = "dodgerblue4"),
    axis.title.y.left  = element_text(color = "green4")
  )
```

```{r line-graph-avg-median-cum with IQR shading for interpretation}
line_df <- thin_film_df %>%
  filter(is.finite(instYr), is.finite(capMW_total)) %>%
  group_by(instYr) %>%
  summarise(
    median_cap   = median(capMW_total, na.rm = TRUE),
    q1_cap       = quantile(capMW_total, 0.25, na.rm = TRUE),
    q3_cap       = quantile(capMW_total, 0.75, na.rm = TRUE),
    annual_count = n(),
    .groups = "drop"
  ) %>%
  arrange(instYr) %>%
  mutate(cumulative_count = cumsum(annual_count))

scale_factor <- max(line_df$q3_cap) / max(line_df$cumulative_count)

ggplot(line_df, aes(x = instYr)) +
  geom_ribbon(aes(ymin = q1_cap, ymax = q3_cap),
              fill = "chartreuse3", alpha = 0.2) +
  geom_line(aes(y = median_cap, color = "Median Capacity (MW)"), linewidth = 1.2) +
  geom_point(aes(y = median_cap, color = "Median Capacity (MW)"), size = 3.5, shape = 18) +
  geom_line(aes(y = cumulative_count * scale_factor, color = "Cumulative Installations"), linewidth = 1.2) +
  geom_point(aes(y = cumulative_count * scale_factor, color = "Cumulative Installations"), size = 2.5) +
  scale_y_continuous(
    name     = "Capacity per Array (MW)",
    sec.axis = sec_axis(~ . / scale_factor, name = "Cumulative Array Count")
  ) +
  scale_x_continuous(breaks = seq(min(line_df$instYr), max(line_df$instYr), by = 2)) +
  scale_color_manual(
    values = c("Median Capacity (MW)"     = "chartreuse3",
               "Cumulative Installations" = "dodgerblue4"),
    name   = NULL
  ) +
  labs(
    title    = "Median Array Capacity Against Cumulative Installations",
    subtitle = "Median capacity per array (IQR shaded) vs cumulative array count by year",
    x        = "Installation Year"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title         = element_text(face = "bold", size = 13),
    plot.subtitle      = element_text(size = 10, color = "grey40"),
    legend.position    = "top",
    panel.grid.minor   = element_blank(),
    axis.title.y.right = element_text(color = "dodgerblue4"),
    axis.title.y.left  = element_text(color = "chartreuse3")
  )
```

```{r line graph with IQR shading - adding outliers}
# Summarise capacity distribution and installation counts per year
# min/max capture the full range; IQR (q1–q3) shows where most arrays fall
line_df <- thin_film_df %>%
  filter(is.finite(instYr), is.finite(capMW_total)) %>%
  group_by(instYr) %>%
  summarise(
    median_cap   = median(capMW_total, na.rm = TRUE),
    q1_cap       = quantile(capMW_total, 0.25, na.rm = TRUE),
    q3_cap       = quantile(capMW_total, 0.75, na.rm = TRUE),
    min_cap      = min(capMW_total,      na.rm = TRUE),
    max_cap      = max(capMW_total,      na.rm = TRUE),
    annual_count = n(),
    .groups      = "drop"
  ) %>%
  arrange(instYr) %>%
  mutate(cumulative_count = cumsum(annual_count))

# Scale factor aligns the two y-axes so max cumulative count maps to max capacity
# This lets both series share the same plot area without distorting either axis
scale_factor <- max(line_df$max_cap) / max(line_df$cumulative_count)

ggplot(line_df, aes(x = instYr)) +
  # IQR ribbon shows the middle 50% of array capacities per year
  geom_ribbon(aes(ymin = q1_cap, ymax = q3_cap),
              fill = "chartreuse3", alpha = 0.2) +
  geom_line(aes(y = median_cap,  color = "Median Capacity (MW)"),    linewidth = 1.2) +
  geom_point(aes(y = median_cap, color = "Median Capacity (MW)"),    size = 3.5, shape = 18) +
  # Triangle points mark the annual min and max — same color, same legend entry
  geom_point(aes(y = min_cap,    color = "Min/Max Capacity (MW)"),   size = 2.5, shape = 17) +
  geom_point(aes(y = max_cap,    color = "Min/Max Capacity (MW)"),   size = 2.5, shape = 17) +
  # Cumulative count scaled to left axis units via scale_factor for dual-axis alignment
  geom_line(aes(y = cumulative_count * scale_factor,  color = "Cumulative Installations"), linewidth = 1.2) +
  geom_point(aes(y = cumulative_count * scale_factor, color = "Cumulative Installations"), size = 2.5) +
  scale_y_continuous(
    name     = "Capacity per Array (MW)",
    sec.axis = sec_axis(~ . / scale_factor, name = "Cumulative Array Count")  # inverse transform recovers true count
  ) +
  scale_x_continuous(breaks = seq(min(line_df$instYr), max(line_df$instYr), by = 2)) +
  scale_color_manual(
    values = c("Median Capacity (MW)"     = "chartreuse3",
               "Cumulative Installations" = "dodgerblue4",
               "Min/Max Capacity (MW)"    = "darkorange"),
    name   = NULL
  ) +
  labs(
    title    = "Median Array Capacity Against Cumulative Installations",
    subtitle = "Median capacity per array (IQR shaded, min/max points) vs cumulative array count by year",
    x        = "Installation Year"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title         = element_text(face = "bold", size = 13),
    plot.subtitle      = element_text(size = 10, color = "grey40"),
    legend.position    = "top",
    panel.grid.minor   = element_blank(),
    axis.title.y.right = element_text(color = "dodgerblue4"),  # right axis color matches cumulative line
    axis.title.y.left  = element_text(color = "chartreuse3")   # left axis color matches capacity line
  )
```
```{r IQR Table of Capacity per Array based on Year}
#install.packages("skimr")
library(skimr)

thin_film_df %>%
  filter(is.finite(instYr)) %>%
  group_by(instYr) %>%
  skim(capMW_total)
```


```{r Investigating the 2014 Outlier}
thin_film_df %>% filter(instYr == 2014)
# 3571 topaz solar farms 
# capMWDC = 752.1 
# capMWAC = 585.9
# Source = GMSEUSgeoref_v2_0
# Est Num Mod: 10,439,497
# Google Num Mod: 8,500,000 
# Google: 550 MW             - Google source https://www.bherenewables.com/topaz-solar-farms
# Google: operational year 2015
# Series: 3
```


```{r cumulative-cog-national}

# Plots the cumulative national center of gravity (CoG) migration over time
# CoG = weighted mean lat/lon of all arrays installed up to and including each year
# weight_col = NULL gives equal weight to each array (count-based CoG)
# weight_col = a numeric column weights each array by that value (e.g. capacity, material mass)
plot_cog <- function(weight_col = NULL, title_label, xlim = NULL, ylim = NULL) {
  
  # Filter to arrays with valid installation year and coordinates
  d <- thin_film_df %>%
    filter(is.finite(instYr), is.finite(longitude), is.finite(latitude))
  
  # If a weight column is provided, keep only valid positive weights
  if (!is.null(weight_col)) {
    d <- d %>% filter(is.finite(.data[[weight_col]]), .data[[weight_col]] > 0)
  }

  # Get all installation years represented in the filtered dataset
  all_years <- sort(unique(d$instYr))

  # For each year, compute the cumulative weighted mean lat/lon
  # Cumulative means all arrays installed up to and including that year are included
  # This tracks how the national geographic center of mass shifts as deployment grows
  cog_df <- purrr::map_dfr(all_years, function(yr) {
    d_cum <- d %>% filter(instYr <= yr)
    if (nrow(d_cum) == 0) return(NULL)
    
    # Use equal weights if no weight column is supplied;
    # otherwise use the selected numeric weighting variable
    w <- if (is.null(weight_col)) rep(1, nrow(d_cum)) else d_cum[[weight_col]]
    
    tibble(
      instYr  = yr,
      cog_lon = weighted.mean(d_cum$longitude, w = w, na.rm = TRUE),
      cog_lat = weighted.mean(d_cum$latitude,  w = w, na.rm = TRUE),
      total_w = sum(w, na.rm = TRUE)
    )
  }) %>%
    # Remove invalid CoG points and years with zero total weight
    filter(is.finite(cog_lon), is.finite(cog_lat), total_w > 0) %>%
    arrange(instYr)

  # Reproject CoG points to EPSG:5070 and extract x/y coordinates for plotting
  # label_yr labels even years only to avoid overcrowding
  cog_sf <- st_as_sf(cog_df, coords = c("cog_lon", "cog_lat"), crs = 4326) %>%
    st_transform("EPSG:5070") %>%
    mutate(
      x        = st_coordinates(.)[, 1],
      y        = st_coordinates(.)[, 2],
      label_yr = ifelse(instYr %% 3 == 0 | instYr == min(instYr) | instYr == max(instYr), instYr, NA)
    )

  # Build year-to-year CoG segments so the migration path can be colored by year
  # yr_mid is the midpoint year between consecutive points — used to color each segment
  cog_segments <- cog_sf %>%
    st_drop_geometry() %>%
    arrange(instYr) %>%
    mutate(
      x_end  = lead(x),
      y_end  = lead(y),
      yr_mid = (instYr + lead(instYr)) / 2
    ) %>%
    filter(!is.na(x_end))

  # Build the national CoG migration map
  ggplot() +
    geom_sf(data = states_p, fill = "grey75", color = "black", linewidth = 0.15) +
    geom_sf(data = us_p,     fill = NA,       color = "black", linewidth = 0.2) +
    # Segments connect consecutive CoG positions; colored by midpoint year
    geom_segment(data = cog_segments,
                 aes(x = x, y = y, xend = x_end, yend = y_end, color = yr_mid),
                 linewidth = 1.2) +
    geom_point(data = cog_sf, aes(x = x, y = y, color = instYr), size = 2) +
    # Labels on even years only to reduce overplotting
    ggrepel::geom_text_repel(
  data        = cog_sf %>% filter(!is.na(label_yr)),
  aes(x = x, y = y, label = label_yr),
  size        = 2.8,
  fontface    = "bold",
  color       = "black",
  box.padding = 0.5,
  nudge_y = 15000,
  min.segment.length = 0.2  # draws a line from label to point when pushed far
) +
    scale_color_viridis_c(
  option = "plasma",
  name   = "Year",
  limits = c(2006, max(thin_film_df$instYr, na.rm = TRUE)),
  breaks = seq(2006, max(thin_film_df$instYr, na.rm = TRUE), by = 2)
) +
    coord_sf(
  crs    = "EPSG:5070",
  xlim   = if (!is.null(xlim)) xlim else c(min(cog_sf$x) - 400000, max(cog_sf$x) + 400000),
  ylim   = if (!is.null(ylim)) ylim else c(min(cog_sf$y) - 400000, max(cog_sf$y) + 400000),
  expand = FALSE
) +
    theme_void() +
    theme(
      plot.title      = element_text(face = "bold", size = 13),
      plot.subtitle   = element_text(size = 10, color = "grey40"),
      legend.position = "right"
    ) +
    labs(
      title = paste("Cumulative Center of Gravity Migration:", title_label)
    )
}
``` 

```{r Using the Center of gravity functions }
# Call it the same way
plot_cog(NULL,                "Array Installations")
plot_cog("capMW_total",       "Installed Capacity")
plot_cog("total_CdTe_g",      "Cadmium Telluride")
plot_cog("total_Aluminum_kg", "Aluminum")
plot_cog("total_Steel_kg",    "Steel")
plot_cog("total_Copper_g",    "Copper")
plot_cog("total_EVA_kg",      "EVA Encapsulant")
plot_cog("total_Array_Weight_kg", "Total Array Mass")
plot_cog("total_GlassMass_kg", "Glass Mass")
plot_cog("total_CalcMass_kg",  "Total Calculated Mass")
```
```{r}
thin_film_df %>% filter(instYr == 2025) %>% nrow()
```


```{r CoG region bounds and subplots for Manusript}
library("ggrepel")
# Compute shared bounding box across all 7 weighting variables so that the region box is uniform across all subplots
all_cog_points <- purrr::map_dfr(
  list(
    list(col = "total_CalcMass_kg",  name = "Total Calculated Mass"),
    list(col = "total_CdTe_g",       name = "Cadmium Telluride"),
    list(col = "total_Copper_g",     name = "Copper"),
    list(col = "total_Aluminum_kg",  name = "Aluminum"),
    list(col = "total_Steel_kg",     name = "Steel"),
    list(col = "total_GlassMass_kg", name = "Glass Mass"),
    list(col = "total_EVA_kg",       name = "EVA Encapsulant")
  ),
  function(mat) {
    d <- thin_film_df %>%
      filter(is.finite(instYr), is.finite(longitude), is.finite(latitude),
             is.finite(.data[[mat$col]]), .data[[mat$col]] > 0)
    purrr::map_dfr(sort(unique(d$instYr)), function(yr) {
      d_cum <- d %>% filter(instYr <= yr)
      if (nrow(d_cum) == 0) return(NULL)
      tibble(
        cog_lon = weighted.mean(d_cum$longitude, w = d_cum[[mat$col]], na.rm = TRUE),
        cog_lat = weighted.mean(d_cum$latitude,  w = d_cum[[mat$col]], na.rm = TRUE)
      )
    })
  }
) %>%
  st_as_sf(coords = c("cog_lon", "cog_lat"), crs = 4326) %>%
  st_transform("EPSG:5070") %>%
  mutate(
    x = st_coordinates(.)[, 1],
    y = st_coordinates(.)[, 2]
  )

pad <- 150000   #  so that there isnt any CoG points right on the border of the subplots 
shared_xlim <- c(min(all_cog_points$x) - pad, max(all_cog_points$x) + pad)
shared_ylim <- c(min(all_cog_points$y) - pad, max(all_cog_points$y) + pad)

# Compute CoG map dimensions and match aspect ratio for region map 
# so the first row of subplots is uniform in size compared to the rest
cog_width    <- shared_xlim[2] - shared_xlim[1]
cog_height   <- shared_ylim[2] - shared_ylim[1]
aspect_ratio <- cog_width / cog_height

us_bbox    <- st_bbox(us_p)
us_height  <- as.numeric(us_bbox["ymax"] - us_bbox["ymin"])
us_width   <- as.numeric(us_bbox["xmax"] - us_bbox["xmin"])
target_width <- us_height * aspect_ratio
x_pad <- (target_width - us_width) / 2

# Centroid Region box map
p_locator <- ggplot() +
  geom_sf(data = states_p, fill = "grey85", color = "black", linewidth = 0.15) +
  geom_sf(data = thin_film_sf, color = "grey25", size = 0.55, alpha = 0.6) +
  geom_sf(data = us_p,     fill = NA,       color = "black", linewidth = 0.2) +
  annotate("rect",
    xmin = shared_xlim[1], xmax = shared_xlim[2],
    ymin = shared_ylim[1], ymax = shared_ylim[2],
    color = "red", fill = NA, linewidth = 1  # zoom region 
  ) +
  coord_sf(
    crs    = "EPSG:5070",
    xlim   = c(us_bbox["xmin"] - x_pad, us_bbox["xmax"] + x_pad),
    expand = FALSE
  ) +
  theme_void() +
  theme(plot.title = element_text(size = 9, face = "bold", hjust = 0)) +
  labs(title = "A.                              Centroid Region")

# Helper to build each subplot with shared zoom and no legend- legend will be added later 
clean_cog <- function(weight_col, title_label) {
  plot_cog(weight_col, title_label, xlim = shared_xlim, ylim = shared_ylim) +
    labs(title = title_label, subtitle = NULL) +
    theme(
      plot.title      = element_text(size = 9, face = "bold", hjust = 0),
      legend.position = "none"
    )
}

# Total Calculated Mass plot 
p_calcmass <- plot_cog("total_CalcMass_kg", "Total Calculated Mass",
                        xlim = shared_xlim, ylim = shared_ylim) +
  labs(title = "B.                         Total Calculated Mass", subtitle = NULL) +
  theme(
    plot.title      = element_text(size = 9, face = "bold", hjust = 0),
    legend.position = "none"
  )

# Individual material subplots we selected, can be changed out 
p_cdte   <- clean_cog("total_CdTe_g",        "C.                                Cadmium Telluride")
p_copper <- clean_cog("total_Copper_g",      "E.                                      Copper")
p_al     <- clean_cog("total_Aluminum_kg",   "G.                                      Aluminum")
p_steel  <- clean_cog("total_Steel_kg",      "F.                                       Steel")
p_glass  <- clean_cog("total_GlassMass_kg",  "D.                                    Glass Mass")
p_eva    <- clean_cog("total_EVA_kg",        "F.                                EVA Encapsulant")

# Shared legend added here
shared_legend <- cowplot::get_legend(
  plot_cog("total_CalcMass_kg", "", xlim = shared_xlim, ylim = shared_ylim) +
    theme(
      legend.position   = "right",
      legend.title      = element_text(size = 9, face = "bold"),
      legend.key.height = unit(1.5, "cm")
    )
)

# layout of the 8 subplots
main_plot <- (p_locator | p_calcmass) /
  (p_cdte    | p_glass) /
  (p_copper      | p_eva)  /
  (p_al   | p_steel)    +
  plot_layout(
    guides  = "collect",
    widths  = c(1, 1)
  ) &
  theme(
    legend.position   = "right",   # so it renders correctly when downloaded 
    legend.key.height = unit(2.5, "cm"),
    legend.title      = element_text(size = 9, face = "bold"),
    legend.text       = element_text(size = 8)
  )
main_plot

ggsave(   # saved locally to import to manuscript 
  filename = "C:/Users/miada/Downloads/Energy_Lab/CoG_Migration_by_Material.png",
  plot     = main_plot,
  width    = 7.5,
  height   = 6,
  units    = "in",
  dpi      = 300
)
```
```{r}
thin_film_df <- array_based_df %>%
  filter(modType == "cdte") %>%
  ...
```


```{r Checking Source of MW estimations to determine reliablility in the future analysis}
thin_film_df %>%
  group_by(Source) %>%
  summarise(n_arrays = n(), .groups = "drop") %>%
  mutate(pct = round(n_arrays / 456 * 100, 1)) %>%
  arrange(desc(n_arrays)) %>%
  print(n = Inf)

```



### Results Metircs
```{r Nationwide stats}
# Total modules, capacity, and arrays nationwide
thin_film_df %>%
  summarise(
    total_arrays   = n(),
    total_modules  = sum(num_modules, na.rm = TRUE),
    total_cap_MW   = sum(capMW_total, na.rm = TRUE),
    total_CdTe_t   = sum(total_CdTe_g, na.rm = TRUE) / 1e6,
    total_Al_t     = sum(total_Aluminum_kg, na.rm = TRUE) / 1e3,
    total_Steel_t  = sum(total_Steel_kg, na.rm = TRUE) / 1e3,
    total_EVA_t    = sum(total_EVA_kg, na.rm = TRUE) / 1e3,
    total_Copper_t = sum(total_Copper_g, na.rm = TRUE) / 1e6,
    total_Glass_t     = sum(total_GlassMass_kg, na.rm = TRUE) / 1e3,
    total_CalcMass_t  = sum(total_CalcMass_kg,  na.rm = TRUE) / 1e3
  )
```


```{r By series summary}
thin_film_df %>%
  group_by(predicted_series) %>%
  summarise(
    num_arrays        = n(),
    total_modules     = sum(num_modules,        na.rm = TRUE),
    total_cap_MW      = sum(capMW_total,         na.rm = TRUE),
    total_CdTe_t      = sum(total_CdTe_g,        na.rm = TRUE) / 1e6,
    total_Al_t        = sum(total_Aluminum_kg,   na.rm = TRUE) / 1e3,
    total_Steel_t     = sum(total_Steel_kg,      na.rm = TRUE) / 1e3,
    total_EVA_t       = sum(total_EVA_kg,        na.rm = TRUE) / 1e3,
    total_Copper_t    = sum(total_Copper_g,      na.rm = TRUE) / 1e6,
    total_Glass_t     = sum(total_GlassMass_kg,  na.rm = TRUE) / 1e3,
    total_CalcMass_t  = sum(total_CalcMass_kg,   na.rm = TRUE) / 1e3,
    .groups = "drop"
  ) %>%
  mutate(
    pct_arrays    = round(100 * num_arrays    / sum(num_arrays),    1),
    pct_modules   = round(100 * total_modules / sum(total_modules), 1),
    pct_cap       = round(100 * total_cap_MW  / sum(total_cap_MW),  1),
    pct_Glass     = round(100 * total_Glass_t / sum(total_Glass_t), 1),
    pct_CalcMass  = round(100 * total_CalcMass_t / sum(total_CalcMass_t), 1)
  ) %>%
  arrange(predicted_series)
```

```{r By Series: Material Intensity (kg per MW)}
thin_film_df %>%
  group_by(predicted_series) %>%
  summarise(
    total_cap_MW     = sum(capMW_total,        na.rm = TRUE),
    total_CdTe_kg    = sum(total_CdTe_g,       na.rm = TRUE) / 1e3,   # g → kg
    total_Al_kg      = sum(total_Aluminum_kg,  na.rm = TRUE),
    total_Steel_kg   = sum(total_Steel_kg,     na.rm = TRUE),
    total_Copper_kg  = sum(total_Copper_g,     na.rm = TRUE) / 1e3,   # g → kg
    total_Glass_kg   = sum(total_GlassMass_kg, na.rm = TRUE),
    total_CalcMass_kg = sum(total_CalcMass_kg, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    CdTe_kg_per_MW     = round(total_CdTe_kg     / total_cap_MW, 2),
    Al_kg_per_MW       = round(total_Al_kg        / total_cap_MW, 2),
    Steel_kg_per_MW    = round(total_Steel_kg     / total_cap_MW, 2),
    Copper_kg_per_MW   = round(total_Copper_kg    / total_cap_MW, 2),
    Glass_kg_per_MW    = round(total_Glass_kg     / total_cap_MW, 2),
    CalcMass_kg_per_MW = round(total_CalcMass_kg  / total_cap_MW, 2)
  ) %>%
  select(predicted_series, ends_with("_per_MW"))
```



```{r Top 10 States}
thin_film_df %>%
  group_by(state_name) %>%
  summarise(
    num_arrays        = n(),
    total_modules     = sum(num_modules,        na.rm = TRUE),
    total_cap_MW      = sum(capMW_total,         na.rm = TRUE),
    total_CdTe_t      = sum(total_CdTe_g,        na.rm = TRUE) / 1e6,
    total_Al_t        = sum(total_Aluminum_kg,   na.rm = TRUE) / 1e3,
    total_Steel_t     = sum(total_Steel_kg,      na.rm = TRUE) / 1e3,
    total_EVA_t       = sum(total_EVA_kg,        na.rm = TRUE) / 1e3,
    total_Copper_t    = sum(total_Copper_g,      na.rm = TRUE) / 1e6,
    total_Glass_t     = sum(total_GlassMass_kg,  na.rm = TRUE) / 1e3,
    total_CalcMass_t  = sum(total_CalcMass_kg,   na.rm = TRUE) / 1e3,
    .groups = "drop"
  ) %>%
  mutate(
    pct_modules        = round(100 * total_modules   / sum(total_modules),    1),
    pct_CdTe           = round(100 * total_CdTe_t    / sum(total_CdTe_t),     1),
    pct_Al             = round(100 * total_Al_t       / sum(total_Al_t),       1),
    pct_Steel          = round(100 * total_Steel_t    / sum(total_Steel_t),    1),
    pct_EVA            = round(100 * total_EVA_t      / sum(total_EVA_t),      1),
    pct_Glass          = round(100 * total_Glass_t    / sum(total_Glass_t),    1),
    pct_CalcMass       = round(100 * total_CalcMass_t / sum(total_CalcMass_t), 1),
    CdTe_t_per_MW      = round(total_CdTe_t    / total_cap_MW, 4),
    Al_t_per_MW        = round(total_Al_t       / total_cap_MW, 3),
    Steel_t_per_MW     = round(total_Steel_t    / total_cap_MW, 3),
    EVA_t_per_MW       = round(total_EVA_t      / total_cap_MW, 3),
    Glass_t_per_MW     = round(total_Glass_t    / total_cap_MW, 3),
    CalcMass_t_per_MW  = round(total_CalcMass_t / total_cap_MW, 3)
  ) %>%
  arrange(desc(total_modules)) %>%
  slice_head(n = 10)
```


```{r Array Characteristics by Series}
thin_film_df %>%
  group_by(predicted_series) %>%
  summarise(
    avg_modules_per_array     = mean(num_modules,        na.rm = TRUE),
    avg_cap_MW_per_array      = mean(capMW_total,        na.rm = TRUE),
    median_cap_MW             = median(capMW_total,      na.rm = TRUE),
    avg_area_m2               = mean(totRowArea,         na.rm = TRUE),
    avg_efficiency            = mean(effInit,            na.rm = TRUE),
    avg_cap_MW                = mean(capMW,              na.rm = TRUE),
    avg_Glass_kg_per_array    = mean(total_GlassMass_kg, na.rm = TRUE),
    avg_CalcMass_kg_per_array = mean(total_CalcMass_kg,  na.rm = TRUE),
    .groups = "drop"
  )
```



```{r Distribution of Arrays, Modules, Capacity, and Total Mass by Series}
library(tidyverse)

figure5_data <- thin_film_df %>%
  group_by(predicted_series) %>%
  summarise(
    num_arrays       = n(),
    total_modules    = sum(num_modules,      na.rm = TRUE),
    total_cap_MW     = sum(capMW_total,      na.rm = TRUE),
    total_CalcMass_t = sum(total_CalcMass_kg, na.rm = TRUE) / 1e3,
    .groups = "drop"
  ) %>%
  mutate(
    pct_arrays    = round(100 * num_arrays       / sum(num_arrays),       1),
    pct_modules   = round(100 * total_modules    / sum(total_modules),    1),
    pct_capacity  = round(100 * total_cap_MW     / sum(total_cap_MW),     1),
    pct_mass      = round(100 * total_CalcMass_t / sum(total_CalcMass_t), 1)
  ) %>%
  select(predicted_series, pct_arrays, pct_modules, pct_capacity, pct_mass) %>%
  pivot_longer(
    cols      = c(pct_arrays, pct_modules, pct_capacity, pct_mass),
    names_to  = "metric",
    values_to = "percentage"
  ) %>%
  mutate(
    metric = recode(metric,
      "pct_arrays"   = "Arrays (%)",
      "pct_modules"  = "Modules (%)",
      "pct_capacity" = "Capacity (%)",
      "pct_mass"     = "Total Mass (%)"
    ),
    metric = factor(metric, levels = c("Arrays (%)", "Modules (%)", "Capacity (%)", "Total Mass (%)"))
  )

ggplot(figure5_data, aes(x = percentage, y = predicted_series, fill = metric)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.75,
           color = "black", linewidth = 0.3) +
  geom_text(aes(label = paste0(percentage,"%"), x = percentage),
            position = position_dodge(width = 0.75),
            vjust = 0.43, hjust = 1.1, size = 4, fontface = "bold", color = "white") +
  scale_fill_manual(values = c(
    "Arrays (%)"    = "dodgerblue4",
    "Modules (%)"   = "mediumpurple2",
    "Capacity (%)"  = "forestgreen",
    "Total Mass (%)" = "magenta3"
  )) +
  scale_x_continuous(limits = c(0, 55), breaks = seq(0, 50, 10),
                     expand = expansion(mult = c(0, 0.05))) +
  scale_y_discrete(limits = rev) +
  labs(
    x        = "Percentage of National Total (%)",
    y        = "Module Series",
    fill     = "Legend"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position      = c(0.83, 0.95),  # top-right corner
    legend.justification = c(1, 1),         # anchor point is top-right of legend box
    legend.margin        = margin(4, 8, 4, 8),
    legend.background    = element_rect(fill = "white", color = "grey80", linewidth = 0.3),
    legend.text          = element_text(size = 11),
    legend.title         = element_text(size = 11, face = "bold"),
    legend.key.size      = unit(1.2, "cm"),
    axis.text.x          = element_text(size = 10),
    axis.text.y          = element_text(face = "bold", size = 11),
    panel.grid.major.y   = element_blank(),
    panel.grid.minor     = element_blank(),
    plot.background      = element_rect(fill = "white", color = NA),
    panel.background     = element_rect(fill = "white", color = NA)
  )


ggsave(
  filename = "C:/Users/miada/Downloads/Energy_Lab/figure5_series_distribution.png",
  width    = 15,
  height   = 6,
  units    = "in",
  dpi      = 600
)
```





```{r Material Intensity per Module by Series - Table PNG}
library(knitr)
library(kableExtra)
library(webshot2)
library(magick)

material_per_module_table <- tibble(
  Series                      = c("2", "3", "4", "6", "7"),
  `Module Area (m2)`             = c(0.72,  0.72,  0.72,  2.47,  2.80),
  `CdTe (g/mod)`               = c(13,    13,    11,    36,    41  ),
  `Glass (kg/mod)`             = c(11.52, 11.52, 11.52, 29.08, 29.07),
  `Aluminum (kg/mod)`          = c(NA,    NA,    NA,    3.9,  NA  ),
  `Steel (kg/mod)`             = c(NA,    NA,    NA,    NA,    6.44),
  `EVA (kg/mod)`               = c(0.28,  0.28,  0.28,  0.97,  1.11),
  `Copper (g/mod)`             = c(43.7,  27.3,  27.3,  16.1,  14.6),
  `Other (kg/mod)`             = c(0.30,  0.30,  0.30,  0.30,  0.30),
  `Total Mass (kg/mod)`          = c(12.16, 12.14, 12.14, 34.30, 36.98),
  `Reported Mass (kg/mod)`     = c(12, 12, 12, 34.5, 39.7 ),
  `Diff (%)`                   = round(100 * (c(12.16, 12.14, 12.14, 34.30, 36.98) -
                                              c(12,    12,    12,    34.5,  39.7 )) /
                                              c(12,    12,    12,    34.5,  39.7 ), 1)
) %>%
  mutate(across(where(is.numeric), ~ifelse(is.na(.), "-", as.character(.))))


material_per_module_table %>%
  kbl(align = c("l", rep("c", 9)), booktabs = TRUE) %>%
  kable_styling(
    full_width        = FALSE,
    font_size         = 20,
    bootstrap_options = c("condensed")
  ) %>%
  row_spec(0, bold = TRUE, background = "#18453B", color = "white") %>%
  row_spec(0:5, extra_css = "border-bottom: 1px solid black;") %>%
  column_spec(1,  bold = TRUE, background = "#008934", color = "white", extra_css = "text-align: center;") %>%
  column_spec(2,  extra_css = "border-right: 1px solid black;") %>%
  column_spec(4,  background = "#D6F5E3") %>%                                            # Glass — now light green
  column_spec(10, background = "#0B996E", extra_css = "border-left: 3px solid black;") %>%  # Tot mass — stays dark green
  column_spec(11, width = "2.2cm", extra_css = "text-align: center;") %>% 
  column_spec(12, width = "1.8cm", extra_css = "text-align: center;") %>% 
  column_spec(12, background = "#0B996E") %>%      # Diff (%) — new, matches Tot mas 
  save_kable("C:/Users/miada/Downloads/Energy_Lab/material_intensity_per_module.png",
             vwidth  = 1400,   # bumped up from default 992 to fit 12 columns
    vheight = 500,
    zoom    = 3 
    )
```

```{r}
colnames(thin_film_df)
write.csv(thin_film_df, "C:/Users/miada/Downloads/Energy_Lab/thin_film_df.csv", row.names = FALSE)
```

```{r series4-6-overlap-justification}
## Series 4 vs Series 6 — Overlap Justification Figure
# --- Panel A: Installation year overlap timeline ---
overlap_years <- tibble(
  predicted_series = c("Series 4", "Series 6"),
  start_year       = c(2016, 2018),
  end_year         = c(2019, 2022)
)

overlap_window <- tibble(start = 2018, end = 2019)

p_timeline <- ggplot(overlap_years) +
  geom_rect(data = overlap_window,
            aes(xmin = start, xmax = end, ymin = 0.4, ymax = 2.6),
            fill = "grey85", alpha = 0.6, inherit.aes = FALSE) +
  geom_segment(aes(x = start_year, xend = end_year,
                    y = predicted_series, yend = predicted_series,
                    color = predicted_series),
               linewidth = 8, lineend = "round") +
  geom_text(aes(x = (start_year + end_year) / 2, y = predicted_series,
                label = paste0(start_year, "\u2013", end_year)),
            color = "white", fontface = "bold", size = 3.5) +
  annotate("text", x = 2018.5, y = 2.6,
           label = "Overlap window\n(2018\u20132019)",
           vjust = -0.3, size = 3.2, fontface = "italic", color = "grey30") +
  scale_color_manual(values = c("Series 4" = "#96CDCD", "Series 6" = "navy")) +
  scale_x_continuous(breaks = 2016:2022) +
  labs(title = "Series 4 vs. Series 6 Installation Windows",
       x = "Installation Year", y = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none",
        panel.grid.minor  = element_blank(),
        axis.text.y       = element_text(face = "bold"))

# --- Panel B: Proportion of arrays affected ---
# --- Panel B: Proportion of arrays affected, with overlap-year breakdown ---
overlap_summary <- thin_film_df %>%
  mutate(
    group = case_when(
      predicted_series %in% c("Series 4", "Series 6") & instYr %in% c(2018, 2019) ~
        "Series 4/6 — overlap years (2018–2019)",
      predicted_series %in% c("Series 4", "Series 6") ~
        "Series 4/6 — other years",
      TRUE ~ "Other series"
    ),
    group = factor(group, levels = c(
      "Other series",
      "Series 4/6 — other years",
      "Series 4/6 — overlap years (2018–2019)"
    ))
  ) %>%
  count(group) %>%
  mutate(pct = round(100 * n / sum(n), 1))

p_proportion <- ggplot(overlap_summary, aes(x = "", y = pct, fill = group)) +
  geom_col(width = 0.5, color = "black", linewidth = 0.3) +
  geom_text(aes(label = paste0(pct, "%")),
            position = position_stack(vjust = 0.5),
            color = "white", fontface = "bold", size = 3.6) +
  scale_fill_manual(values = c(
    "Other series"                            = "grey80",
    "Series 4/6 — other years"                = "#96CDCD",
    "Series 4/6 — overlap years (2018–2019)"  = "navy"
  )) +
  labs(title = "Share of Arrays Requiring\nSeries 4/6 Disambiguation",
       x = NULL, y = "% of Total Arrays", fill = NULL) +
  guides(fill = guide_legend(ncol = 1)) +
  theme_minimal(base_size = 12) +
  theme(legend.position    = "bottom",
        legend.text        = element_text(size = 8.5),
        panel.grid.major.x = element_blank())

# --- Combine ---
p_timeline + p_proportion +
  plot_layout(widths = c(1.6, 1)) +
  plot_annotation(
    title    = "Why Series 4 and 6 Classification Needs Refinement",
    theme    = theme(plot.title = element_text(face = "bold", size = 14))
  )
```

```{r}
library(plotly)
library(dplyr)

# --- Node order ---
series_levels   <- c("Series 2", "Series 3", "Series 4", "Series 6", "Series 7")  # top -> bottom
material_levels <- c("Glass", "EVA", "CdTe", "Copper", "Other", "Aluminum", "Steel")  # Steel at very bottom, Aluminum second-to-last
node_labels      <- c(series_levels, material_levels)
node_labels_bold <- paste0("<b>", node_labels, "</b>")

# --- Link colors defined FIRST, before anything references them ---
series_link_colors <- c(
  "Series 2" = "rgba(183,228,160,0.85)",  # extra opacity so it stays visible even when thin
  "Series 3" = "rgba(90,174,97,0.55)",
  "Series 4" = "rgba(69,123,157,0.55)",
  "Series 6" = "rgba(185,131,196,0.55)",
  "Series 7" = "rgba(58,0,96,0.55)"
)

# --- Total module counts per series (used to normalize mass -> mass per module) ---
total_modules_lookup <- tibble(
  predicted_series = c("Series 2", "Series 3", "Series 4", "Series 6", "Series 7"),
  total_modules     = c(5686490, 41009416, 39113605, 29798952, 30527452)
)

# --- Build source/target/value from sankey_long, normalized to mass per module ---
sankey_plotly_df <- sankey_long %>%
  mutate(
    predicted_series = as.character(predicted_series),
    material         = as.character(material)
  ) %>%
  left_join(total_modules_lookup, by = "predicted_series") %>%
  mutate(
    mass_kg_per_module = mass_kg / total_modules,
    source = match(predicted_series, node_labels) - 1,
    target = match(material, node_labels) - 1
  ) %>%
  arrange(desc(mass_kg_per_module))   # largest flows first, smallest drawn LAST = on top

link_colors <- unname(series_link_colors[sankey_plotly_df$predicted_series])

# --- Node colors ---
node_colors <- c(
  unname(c("Series 2"="#B7E4A0","Series 3"="#5AAE61","Series 4"="#457B9D",
           "Series 6"="#B983C4","Series 7"="#3A0060")[series_levels]),
  rep("#4A4A4A", length(material_levels))
)

n_series   <- length(series_levels)
n_material <- length(material_levels)

# --- Compute each node's total flow value (drives node height / y position) ---
series_totals <- sankey_plotly_df %>%
  group_by(predicted_series) %>%
  summarise(total = sum(mass_kg_per_module), .groups = "drop") %>%
  slice(match(series_levels, predicted_series))

material_totals <- sankey_plotly_df %>%
  group_by(material) %>%
  summarise(total = sum(mass_kg_per_module), .groups = "drop") %>%
  slice(match(material_levels, material))

make_y_positions <- function(totals, gap = 0.03) {
  totals <- totals / sum(totals)
  n <- length(totals)
  usable_height <- 1 - gap * (n - 1)
  heights <- totals * usable_height
  y_start <- cumsum(c(0, heights[-n])) + gap * (seq_len(n) - 1)
  y_mid   <- y_start + heights / 2
  pmin(pmax(y_mid, 0.01), 0.99)
}

y_series   <- make_y_positions(series_totals$total)
y_material <- make_y_positions(material_totals$total)

node_x <- c(rep(0.01, n_series), rep(0.99, n_material))
node_y <- c(y_series, y_material)

# --- Build the Sankey ---
p <- plot_ly(
  type   = "sankey",
  orientation = "h",
  arrangement = "fixed",
  node = list(
    label     = node_labels_bold,
    color     = node_colors,
    x         = node_x,
    y         = node_y,
    pad       = 18,
    thickness = 14,
    line      = list(color = "white", width = 0),
    textfont  = list(color = "black", size = 13, family = "Arial")
  ),
  link = list(
    source = sankey_plotly_df$source,
    target = sankey_plotly_df$target,
    value  = sankey_plotly_df$mass_kg_per_module,
    color  = link_colors
  )
) %>%
  layout(
    font = list(size = 13, family = "Arial", color = "black"),
    plot_bgcolor  = "white",
    paper_bgcolor = "white",
    margin = list(t = 70, l = 20, r = 20, b = 20)
  )

p
```
```{r}
library(plotly)
library(dplyr)

# --- Node order for the "lower mass" view ---
material_levels_low <- c("EVA", "CdTe", "Copper", "Other")  # top bottom, adjust order as you like
node_labels_low      <- c(series_levels, material_levels_low)
node_labels_low_bold <- paste0("<b>", node_labels_low, "</b>")

# --- Filter existing sankey_plotly_df down to just these materials ---
sankey_plotly_df_low <- sankey_plotly_df %>%
  filter(material %in% material_levels_low) %>%
  mutate(
    source = match(predicted_series, node_labels_low) - 1,
    target = match(material, node_labels_low) - 1
  ) %>%
  arrange(desc(mass_kg_per_module))

link_colors_low <- unname(series_link_colors[sankey_plotly_df_low$predicted_series])

# --- Node colors ---
node_colors_low <- c(
  unname(c("Series 2"="#B7E4A0","Series 3"="#5AAE61","Series 4"="#457B9D",
           "Series 6"="#B983C4","Series 7"="#3A0060")[series_levels]),
  rep("#4A4A4A", length(material_levels_low))
)

n_series_low   <- length(series_levels)
n_material_low <- length(material_levels_low)

# --- Recompute totals + y-positions for this subset only ---
series_totals_low <- sankey_plotly_df_low %>%
  group_by(predicted_series) %>%
  summarise(total = sum(mass_kg_per_module), .groups = "drop") %>%
  slice(match(series_levels, predicted_series))

material_totals_low <- sankey_plotly_df_low %>%
  group_by(material) %>%
  summarise(total = sum(mass_kg_per_module), .groups = "drop") %>%
  slice(match(material_levels_low, material))

y_series_low   <- make_y_positions(series_totals_low$total)
y_material_low <- make_y_positions(material_totals_low$total)

node_x_low <- c(rep(0.01, n_series_low), rep(0.99, n_material_low))
node_y_low <- c(y_series_low, y_material_low)

# --- Build the "lower mass" Sankey ---
p_low <- plot_ly(
  type   = "sankey",
  orientation = "h",
  arrangement = "fixed",
  node = list(
    label     = node_labels_low_bold,
    color     = node_colors_low,
    x         = node_x_low,
    y         = node_y_low,
    pad       = 18,
    thickness = 14,
    line      = list(color = "white", width = 0),
    textfont  = list(color = "black", size = 13, family = "Arial")
  ),
  link = list(
    source = sankey_plotly_df_low$source,
    target = sankey_plotly_df_low$target,
    value  = sankey_plotly_df_low$mass_kg_per_module,
    color  = link_colors_low
  )
) %>%
  layout(
    font = list(size = 13, family = "Arial", color = "black"),
    plot_bgcolor  = "white",
    paper_bgcolor = "white",
    margin = list(t = 70, l = 20, r = 20, b = 20)
  )

p_low
```


```{r}
sankey_plotly_df %>%
  group_by(predicted_series) %>%
  summarise(total = sum(mass_kg_per_module), n_materials = n())
```



```{r}
# One-time setup for static export (kaleido backend)
# install.packages("reticulate")
# reticulate::install_miniconda()
# reticulate::py_install("kaleido")

save_image(p, "C:/Users/miada/Downloads/Energy_Lab/sankey.png",
           width = 1600, height = 900, scale = 3)

# Vector format if your journal wants true vector (SVG/PDF)
save_image(p, "C:/Users/miada/Downloads/Energy_Lab/sankey_publication.svg",
           width = 1600, height = 900)
save_image(p, "C:/Users/miada/Downloads/Energy_Lab/sankey_publication.pdf",
           width = 1600, height = 900)
```


```{r}
library(plotly)
library(dplyr)

# --- Reuse the same summary math, but keep raw totals (not yet %) for flexibility ---
sankey_metric_long <- thin_film_df %>%
  group_by(predicted_series) %>%
  summarise(
    Arrays   = n(),
    Modules  = sum(num_modules,       na.rm = TRUE),
    Capacity = sum(capMW_total,       na.rm = TRUE),
    `Total Mass` = sum(total_CalcMass_kg, na.rm = TRUE) / 1e3,
    .groups = "drop"
  ) %>%
  pivot_longer(-predicted_series, names_to = "metric", values_to = "raw_value") %>%
  group_by(metric) %>%
  mutate(pct_of_metric = 100 * raw_value / sum(raw_value)) %>%   # % of that metric's national total
  ungroup()

# --- Node order: series top->bottom as before, metrics in fixed order ---
series_levels <- c("Series 7", "Series 6", "Series 4", "Series 3", "Series 2")
metric_levels <- c("Arrays", "Modules", "Capacity", "Total Mass")
node_labels   <- c(series_levels, metric_levels)
node_labels_bold <- paste0("<b>", node_labels, "</b>")

sankey_plotly_df <- sankey_metric_long %>%   # or sankey_long, for the material version
  mutate(
    predicted_series = as.character(predicted_series),
    source = match(predicted_series, node_labels) - 1,
    target = match(metric, node_labels) - 1     # or `material` for the other chart
  ) %>%
  arrange(desc(pct_of_metric))   # <- largest values first, smallest (Series 2) drawn LAST = on top

link_colors <- unname(series_link_colors[sankey_plotly_df$predicted_series])

node_colors <- c(
  unname(c("Series 7"="#3A0060","Series 6"="#B983C4","Series 4"="#457B9D",
           "Series 3"="#5AAE61","Series 2"="#B7E4A0")[series_levels]),
  rep("#4A4A4A", length(metric_levels))
)

series_link_colors <- c(
  "Series 2" = "rgba(183,228,160,0.55)",
  "Series 3" = "rgba(90,174,97,0.55)",
  "Series 4" = "rgba(69,123,157,0.55)",
  "Series 6" = "rgba(185,131,196,0.55)",
  "Series 7" = "rgba(58,0,96,0.55)"
)
link_colors <- unname(series_link_colors[sankey_plotly_df$predicted_series])

n_series <- length(series_levels)
n_metric <- length(metric_levels)

# --- Proportional y positions so nodes never overlap ---
series_totals <- sankey_plotly_df %>%
  group_by(predicted_series) %>%
  summarise(total = sum(pct_of_metric), .groups = "drop") %>%
  slice(match(series_levels, predicted_series))

metric_totals <- sankey_plotly_df %>%
  group_by(metric) %>%
  summarise(total = sum(pct_of_metric), .groups = "drop") %>%
  slice(match(metric_levels, metric))

make_y_positions <- function(totals, gap = 0.03) {
  totals <- totals / sum(totals)
  n <- length(totals)
  usable_height <- 1 - gap * (n - 1)
  heights <- totals * usable_height
  y_start <- cumsum(c(0, heights[-n])) + gap * (seq_len(n) - 1)
  y_mid   <- y_start + heights / 2
  pmin(pmax(y_mid, 0.01), 0.99)
}

y_series <- make_y_positions(series_totals$total)
y_metric <- make_y_positions(metric_totals$total)

node_x <- c(rep(0.01, n_series), rep(0.99, n_metric))
node_y <- c(y_series, y_metric)

p <- plot_ly(
  type   = "sankey",
  orientation = "h",
  arrangement = "fixed",
  node = list(
    label     = node_labels_bold,
    color     = node_colors,
    x         = node_x,
    y         = node_y,
    pad       = 18,
    thickness = 14,
    line      = list(color = "white", width = 0),
    textfont  = list(color = "black", size = 13, family = "Arial")
  ),
  link = list(
    source = sankey_plotly_df$source,
    target = sankey_plotly_df$target,
    value  = sankey_plotly_df$pct_of_metric,
    color  = link_colors
  )
) %>%
  layout(
    font = list(size = 13, family = "Arial", color = "black"),
    plot_bgcolor  = "white",
    paper_bgcolor = "white",
    margin = list(t = 70, l = 20, r = 20, b = 20)
  )

p

# sVING
#save_image(p, "C:/Users/miada/Downloads/Energy_Lab/sankey_series_metrics.png",
   #        width = 1600, height = 900, scale = 3)
```

```{r}
library(plotly)
library(dplyr)

# --- Node order ---
series_levels <- c("Series 2", "Series 3", "Series 4", "Series 6", "Series 7")  # top -> bottom
metric_levels <- c("Arrays", "Modules", "Capacity", "Total Mass")

node_labels      <- c(series_levels, metric_levels)
node_labels_bold <- paste0("<b>", node_labels, "</b>")

# --- Link colors ---
series_link_colors <- c(
  "Series 2" = "rgba(183,228,160,0.85)",
  "Series 3" = "rgba(90,174,97,0.55)",
  "Series 4" = "rgba(69,123,157,0.55)",
  "Series 6" = "rgba(185,131,196,0.55)",
  "Series 7" = "rgba(58,0,96,0.55)"
)

# --- Build source/target/value from the metrics dataset ---
sankey_plotly_df <- sankey_metric_long %>%
  mutate(
    predicted_series = as.character(predicted_series),
    source = match(predicted_series, node_labels) - 1,
    target = match(metric, node_labels) - 1
  ) %>%
  arrange(desc(pct_of_metric))   # largest flows first, Series 2 (thinnest) drawn LAST = on top

link_colors <- unname(series_link_colors[sankey_plotly_df$predicted_series])

# --- Node colors ---
node_colors <- c(
  unname(c("Series 2"="#B7E4A0","Series 3"="#5AAE61","Series 4"="#457B9D",
           "Series 6"="#B983C4","Series 7"="#3A0060")[series_levels]),
  rep("#4A4A4A", length(metric_levels))
)

n_series <- length(series_levels)
n_metric <- length(metric_levels)

# --- Compute each node's total flow value (drives node height / y position) ---
series_totals <- sankey_plotly_df %>%
  group_by(predicted_series) %>%
  summarise(total = sum(pct_of_metric), .groups = "drop") %>%
  slice(match(series_levels, predicted_series))

metric_totals <- sankey_plotly_df %>%
  group_by(metric) %>%
  summarise(total = sum(pct_of_metric), .groups = "drop") %>%
  slice(match(metric_levels, metric))

make_y_positions <- function(totals, gap = 0.03) {
  totals <- totals / sum(totals)
  n <- length(totals)
  usable_height <- 1 - gap * (n - 1)
  heights <- totals * usable_height
  y_start <- cumsum(c(0, heights[-n])) + gap * (seq_len(n) - 1)
  y_mid   <- y_start + heights / 2
  pmin(pmax(y_mid, 0.01), 0.99)
}

y_series <- make_y_positions(series_totals$total)
y_metric <- make_y_positions(metric_totals$total)

node_x <- c(rep(0.01, n_series), rep(0.99, n_metric))
node_y <- c(y_series, y_metric)

# --- Build the Sankey ---
p <- plot_ly(
  type   = "sankey",
  orientation = "h",
  arrangement = "fixed",
  node = list(
    label     = node_labels_bold,
    color     = node_colors,
    x         = node_x,
    y         = node_y,
    pad       = 18,
    thickness = 14,
    line      = list(color = "white", width = 0),
    textfont  = list(color = "black", size = 13, family = "Arial")
  ),
  link = list(
    source = sankey_plotly_df$source,
    target = sankey_plotly_df$target,
    value  = sankey_plotly_df$pct_of_metric,
    color  = link_colors
  )
) %>%
  layout(
    font = list(size = 13, family = "Arial", color = "black"),
    plot_bgcolor  = "white",
    paper_bgcolor = "white",
    margin = list(t = 30, l = 20, r = 20, b = 20)  # smaller top margin now that title is gone
  )

p
```


